---
title: "NC_revision"
author: "Jinshi"
date: "5/26/2021"
output: html_document
Note: before run this Rmd file, need to run "Topdown_GlobalRs.Rmd" first 
---

```{r preliminaries, message=FALSE, echo=FALSE}
# Constants
OUTPUT_DIR <- "outputs/"
DATA_DIR <- "data"
# Create output and log folders if they do not exist
if (!file.exists(OUTPUT_DIR)) dir.create(OUTPUT_DIR)

library(kableExtra)
# Load required packages
library(data.table)
library(lubridate)
library(knitr)
library(ggpubr)

library(ggplot2)
theme_set(theme_bw())
library(cowplot)
library(ggmap)
library(maps)
library(mapdata)
# install.packages("ggExtra")
library(ggExtra)

library(tidyr)
library(hexbin)
library(dplyr)

# install.packages('ggbeeswarm')
library(ggbeeswarm)
library(ggridges)
library(patchwork)

# Source all needed functions
source("functions.R")
source("functions-stats.R")
source("functions-resampling.R")
source("functions-calc_bootstrap.R")

# Set chunks defaults; these options will be applied to all subsequent chunks
knitr::opts_chunk$set(
  results = "hide", message = TRUE, include = TRUE,
  echo = FALSE, warning = FALSE,
  fig.height = 4, fig.width = 8, cache = FALSE
)
```

```{r load data}
# input data
read_file <- function(x) read.csv(file.path(DATA_DIR, x), comment.char = "#", stringsAsFactors = FALSE)

GPP <- read_file("GlobalGPP_Sum.csv") # GPP estimates from published paper
GPP %>% filter(Year >= 1980 & Flag != "Model") -> GPP

GlobalRs <- read_file("GlobalRs.csv") # estimates of global Rs
GlobalRs %>% filter(Pub_year > 1980) -> GlobalRs

# aggregate by year to test the temporal representativeness issue
GPP %>% dplyr::select(Year, GPP, SD) %>% 
  group_by(Year) %>% 
  summarise_at(vars(GPP, SD), list(mean), na.rm=T) -> GPP_agg

GPP %>% dplyr::select(Group2, GPP, SD) %>% 
  group_by(Group2) %>% 
  summarise_at(vars(GPP, SD), list(mean), na.rm=T) -> GPP_group

GPP %>% dplyr::select(Year, Group2, GPP, SD) %>% 
  group_by(Year, Group2) %>% 
  summarise_at(vars(GPP, SD), list(mean), na.rm=T) -> GPP_agg_group

GlobalRs %>% dplyr::select(Pub_year, Rs, SD) %>% 
  group_by(Pub_year) %>% 
  summarise_at(vars(Rs, SD), list(mean), na.rm=T) -> GlobalRs_agg

```

```{r bootstrap Rs_raw and GPP_raw}
## bootstrap and test how data temporal uneven distribution affect global Rs and GPP
set.seed(20210526)
bootstrap_data2 <- tibble(
  Rs_raw = sapply(rep(sample_size, N_SAMPLES), RESAMPLE, GlobalRs$Rs, GlobalRs$SD),
  Rs_raw_agg = sapply(rep(sample_size, N_SAMPLES), RESAMPLE, GlobalRs_agg$Rs, GlobalRs_agg$SD),
  GPP_raw = sapply(rep(sample_size, N_SAMPLES), RESAMPLE, GPP$GPP, GPP$SD),
  GPP_raw_agg = sapply(rep(sample_size, N_SAMPLES), RESAMPLE, GPP_agg$GPP, GPP_agg$SD),
  GPP_raw_group = sapply(rep(sample_size, N_SAMPLES), RESAMPLE, GPP_group$GPP, GPP_group$SD),
)

bootstrap_data2$Rs_raw_agg %>% mean()
bootstrap_data2$GPP_raw_agg %>% mean()
```

```{r plot and show representiveness analysis results}
## rs_results is from "Topdown_GlobalRs.Rmd"
## plot and show how data temporal uneven distribution affect global Rs and GPP
bind_rows(
  rs_results %>% select(`value` = Rs_topdown2) %>% 
  mutate(Rs_type = "Rs_gpp"),
  bootstrap_data2 %>%
  select(`Rs` = Rs_raw, `Rs[agg]` = Rs_raw_agg) %>%
  gather(Rs_type)) %>% 
  ggplot(aes(value, fill = Rs_type)) +
  theme_cowplot() +
  geom_density(stat = "density", alpha = 0.65) +
  theme(legend.position = c(0.75, 0.50)) +
  coord_cartesian(xlim = c(50, 110)) +
  scale_fill_manual("",
    labels = c(expression(Rs[lit]), expression(Rs[lit-agg]), expression(Rs[GPP])),
    values=c("skyblue", "blue", "#E41A1C")
  ) +
  theme(
    legend.title = element_blank(),
    legend.text.align = 0
  ) +
  ylab("Density") +
  xlab(expression(Rs ~ (Pg ~ C ~ yr^{-1}))) -> Rs_lit_plot

## gpp_results is from "Topdown_GlobalRs.Rmd"
bind_rows(
  gpp_results %>% select(`value` = GPP2) %>% 
  mutate(GPP_type = "GPP_Rs"),
  bootstrap_data2 %>%
  select(`GPP` = GPP_raw, `GPP[agg]` = GPP_raw_agg) %>%
  gather(GPP_type)) %>% 
  ggplot(aes(value, fill = GPP_type)) +
  theme_cowplot() +
  geom_density(stat = "density", alpha = 0.65) +
  theme(legend.position = c(0.70, 0.50)) +
  coord_cartesian(xlim = c(80, 210)) +
  scale_fill_manual("",
    labels = c(expression(GPP[lit]), expression(GPP[lit-agg]), expression(GPP[Rs])),
    values=c("#E41A1C", "darkred", "skyblue")
  ) +
  theme(
    legend.title = element_blank(),
    legend.text.align = 0
  ) +
  ylab("Density") +
  xlab(expression(GPP ~ (Pg ~ C ~ yr^{-1}))) -> GPP_lit_plot
  
```


```{r gpp and rs aggregate by year}
## plot rs and gpp vs year (both raw data and aggregated by year)
p_gpp <- ggplot(GPP, aes(Year, GPP)) +
  geom_point() +
  geom_errorbar(aes(ymin = GPP - SD, ymax = GPP + SD), na.rm = TRUE) +
  geom_smooth(method = "lm") +
  theme_cowplot() +
  xlim(1980, 2020) +
  ylab(expression(GPP ~ (Pg ~ C ~ yr^-1)))

p_rs <- ggplot(GlobalRs, aes(Pub_year, Rs)) +
  geom_point() +
  geom_errorbar(aes(ymin = Rs - SD, ymax = Rs + SD), na.rm = TRUE) +
  geom_smooth(method = "lm") +
  theme_cowplot() +
  xlab("Year") +
  xlim(1980, 2020) +
  ylab(expression(R[S] ~ (Pg ~ C ~ yr^-1)))

p_gpp_agg <- ggplot(GPP_agg, aes(Year, GPP)) +
  geom_point() +
  geom_errorbar(aes(ymin = GPP - SD, ymax = GPP + SD), na.rm = TRUE) +
  geom_smooth(method = "lm") +
  theme_cowplot() +
  xlim(1980, 2020) +
  ylab(expression(GPP ~ (Pg ~ C ~ yr^-1)))

p_rs_agg <- ggplot(GlobalRs_agg, aes(Pub_year, Rs)) +
  geom_point() +
  geom_errorbar(aes(ymin = Rs - SD, ymax = Rs + SD), na.rm = TRUE) +
  geom_smooth(method = "lm") +
  theme_cowplot() +
  xlab("Year") +
  xlim(1980, 2020) +
  ylab(expression(R[S] ~ (Pg ~ C ~ yr^-1)))

```
```{r plot representativeness results, fig.height=9, fig.width=8}
(p_gpp | p_rs) / (p_gpp_agg | p_rs_agg) / (GPP_lit_plot | Rs_lit_plot) +
  plot_annotation(tag_levels = 'a')

# ggsave("outputs/FigureS7.png", width = 9, height = 8)
```


```{r ratio sites spatial coverage plot, fig.height=8, fig.width=9}
# plot and show the spatial coverage of ratio data compared with MAT and MAP

UDel_climate <- read_file("UDel_summarized_climate.csv")

left_join(Froot %>% mutate(Latitude = round(Latitude*2)/2+0.25, 
                 Longitude = round(Longitude*2)/2+0.25),
          UDel_climate,
          by = c("Latitude", "Longitude")) -> Froot_del

left_join(sub_srdb %>% mutate(Latitude = round(Latitude*2)/2+0.25, 
                 Longitude = round(Longitude*2)/2+0.25),
          UDel_climate,
          by = c("Latitude", "Longitude")) -> sub_srdb_del


left_join(
  srdb_v4 %>%
    select(Ecosystem_type, Leaf_habit, Latitude, Longitude, Rs_annual, Ra_annual, Rh_annual, ER, GPP, NPP) %>%
    filter(!is.na(GPP) | !is.na(NPP) | !is.na(ER)) %>%
    mutate(
      Latitude = round(Latitude*2)/2+0.25, 
      Longitude = round(Longitude*2)/2+0.25,
      ER = if_else(is.na(ER), GPP, ER), # assume NEP very small
      RaGPP_ratio = (ER - NPP) / GPP) %>%
    filter(RaGPP_ratio > 0, RaGPP_ratio < 1, !is.na(Ecosystem_type)),
  UDel_climate,
  by = c("Latitude", "Longitude")) -> sub_raGPP_del

## plot Rroot:Ra representative
bind_rows(
  Froot_del %>% select(MAT, MAP) %>% mutate(Source = "Froot"),
  UDel_climate %>% select(MAT, MAP) %>% mutate(Source = "Global")) %>% 
  ggplot(aes(x = MAT, fill = factor(Source))) +
  theme(legend.position = c(0.25,0.75)) +
  # geom_violin(draw_quantiles = c(0.5))
  geom_density(alpha = 0.5) +
  scale_fill_manual(breaks = c("Froot", "Global"),
                    labels = c(expression(R[root]~':'~R[A]), expression(Global)),
                    values = c("chocolate", "coral4"))+
  theme(
    legend.title = element_blank(),
    legend.text.align = 0)+
  labs(x=element_blank(), y=expression(Density)) ->
  rep_rroot_ra_mat

bind_rows(
  Froot_del %>% select(MAT, MAP) %>% mutate(Source = "Froot"),
  UDel_climate %>% select(MAT, MAP) %>% mutate(Source = "Global")) %>% 
  ggplot(aes(x = MAP, fill = factor(Source))) +
  theme(legend.position = c(0.75,0.75)) +
  # geom_violin(draw_quantiles = c(0.5))
  geom_density(alpha = 0.5) +
  scale_fill_manual(breaks = c("Froot", "Global"),
                    labels = c(expression(R[root]~':'~R[A]), expression(Global)),
                    values = c("skyblue", "blue"))+
  xlim(c(0,5000))+
  theme(
    legend.title = element_blank(),
    legend.text.align = 0)+
  labs(x=element_blank(), y=expression(Density)) ->
  rep_rroot_ra_map

## plot rroot:rs representative
bind_rows(
  sub_srdb_del %>% select(MAT, MAP) %>% mutate(Source = "RrRs"),
  UDel_climate %>% select(MAT, MAP) %>% mutate(Source = "Global")) %>% 
  ggplot(aes(x = MAT, fill = factor(Source))) +
  theme(legend.position = c(0.25,0.75)) +
  # geom_violin(draw_quantiles = c(0.5))
  geom_density(alpha = 0.5) +
  scale_fill_manual(breaks = c("RrRs", "Global"),
                    labels = c(expression(R[root]~':'~R[S]), expression(Global)),
                    values = c("chocolate", "coral4"))+
  theme(
    legend.title = element_blank(),
    legend.text.align = 0)+
  labs(x=element_blank(), y=expression(Density)) ->
  rep_rroot_rs_mat

bind_rows(
  sub_srdb_del %>% select(MAT, MAP) %>% mutate(Source = "RrRs"),
  UDel_climate %>% select(MAT, MAP) %>% mutate(Source = "Global")) %>% 
  ggplot(aes(x = MAP, fill = factor(Source))) +
  theme(legend.position = c(0.75,0.75)) +
  # geom_violin(draw_quantiles = c(0.5))
  geom_density(alpha = 0.5) +
  scale_fill_manual(breaks = c("RrRs", "Global"),
                    labels = c(expression(R[root]~':'~R[S]), expression(Global)),
                    values = c("skyblue", "blue"))+
  xlim(c(0,5000)) +
  theme(
    legend.title = element_blank(),
    legend.text.align = 0)+
  labs(x=element_blank(), y=expression(Density)) ->
  rep_rroot_rs_map

## plot ra:GPP representative
bind_rows(
  sub_raGPP_del %>% select(MAT, MAP) %>% mutate(Source = "RaGPP"),
  UDel_climate %>% select(MAT, MAP) %>% mutate(Source = "Global")) %>% 
  ggplot(aes(x = MAT, fill = factor(Source))) +
  theme(legend.position = c(0.25,0.75)) +
  # geom_violin(draw_quantiles = c(0.5))
  geom_density(alpha = 0.5) +
  scale_fill_manual(breaks = c("RaGPP", "Global"),
                    labels = c(expression(R[A]~':'~GPP), expression(Global)),
                    values = c("chocolate", "coral4"))+
  theme(
    legend.title = element_blank(),
    legend.text.align = 0)+
  labs(x=expression(MAT~"("~degree~C~")"), y=expression(Density)) ->
  rep_ra_gpp_mat

bind_rows(
  sub_raGPP_del %>% select(MAT, MAP) %>% mutate(Source = "RaGPP"),
  UDel_climate %>% select(MAT, MAP) %>% mutate(Source = "Global")) %>% 
  ggplot(aes(x = MAP, fill = factor(Source))) +
  theme(legend.position = c(0.75,0.75)) +
  # geom_violin(draw_quantiles = c(0.5))
  geom_density(alpha = 0.5) +
  scale_fill_manual(breaks = c("RaGPP", "Global"),
                    labels = c(expression(R[A]~':'~GPP), expression(Global)),
                    values = c("skyblue", "blue"))+
  xlim(c(0,5000)) +
  theme(
    legend.title = element_blank(),
    legend.text.align = 0)+
  labs(x=expression(MAP~"("~mm~yr^{-1}~")"), y=expression(Density)) ->
  rep_ra_gpp_map


(rep_rroot_rs_mat | rep_rroot_rs_map) / (rep_rroot_ra_mat | rep_rroot_ra_map) / (rep_ra_gpp_mat | rep_ra_gpp_map) +
  plot_annotation(tag_levels = 'a')

# ggsave("outputs/FigureSX.png", width = 9, height = 8)

```

```{r ratio sites spatial coverage plot using fluxcom gpp, fig.height=8, fig.width=8}
# plot and show the spatial coverage of ratio data compared with MAT and MAP

Fluxcon_gpp <- read_file("Fluxcom_summarized_GPP.csv")

left_join(Froot %>% mutate(Latitude = round(Latitude*2)/2+0.25, 
                 Longitude = round(Longitude*2)/2+0.25),
          Fluxcon_gpp,
          by = c("Latitude", "Longitude")) -> Froot_fluxcom

left_join(sub_srdb %>% mutate(Latitude = round(Latitude*2)/2+0.25, 
                 Longitude = round(Longitude*2)/2+0.25),
          Fluxcon_gpp,
          by = c("Latitude", "Longitude")) -> sub_srdb_fluxcom


left_join(
  srdb_v4 %>%
    select(Ecosystem_type, Leaf_habit, Latitude, Longitude, Rs_annual, Ra_annual, Rh_annual, ER, GPP, NPP) %>%
    filter(!is.na(GPP) | !is.na(NPP) | !is.na(ER)) %>%
    mutate(
      Latitude = round(Latitude*2)/2+0.25, 
      Longitude = round(Longitude*2)/2+0.25,
      ER = if_else(is.na(ER), GPP, ER), # assume NEP very small
      RaGPP_ratio = (ER - NPP) / GPP) %>%
    filter(RaGPP_ratio > 0, RaGPP_ratio < 1, !is.na(Ecosystem_type)),
  Fluxcon_gpp,
  by = c("Latitude", "Longitude")) -> sub_raGPP_fluxcom

## plot Rroot:Ra representative
bind_rows(
  Froot_fluxcom %>% select(FLUXCOM_GPP) %>% mutate(Source = "Froot"),
  Fluxcon_gpp %>% select(FLUXCOM_GPP) %>% mutate(Source = "Global")) %>% 
  ggplot(aes(x = FLUXCOM_GPP, fill = factor(Source))) +
  theme(legend.position = c(0.75,0.75)) +
  # geom_violin(draw_quantiles = c(0.5))
  geom_density(alpha = 0.5) +
  scale_fill_manual(breaks = c("Froot", "Global"),
                    labels = c(expression(R[root]~':'~R[A]), expression(Global)),
                    values = c("chocolate", "coral4"))+
  theme(
    legend.title = element_blank(),
    legend.text.align = 0)+
  labs(x=element_blank(), y=expression(Density)) ->
  rep_rroot_ra_gpp

## regresstion between ratios and GPP
lm(Froot_fluxcom$Froot ~ Froot_fluxcom$FLUXCOM_GPP) %>% summary()

Froot_fluxcom %>% select(Froot, FLUXCOM_GPP) %>% 
  ggplot(aes(FLUXCOM_GPP, Froot)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme(
    legend.title = element_blank(),
    legend.text.align = 0)+
  labs(x=element_blank(), y=expression(R[root]~':'~R[A])) ->
  rep_froot_regression

## how this relationship affect GPPrs

ggplot(rs_results, aes(Froot2, Rs_topdown)) +
  geom_hex() +
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") +
  # xlab(expression(R[root]~':'~R[S])) +
  # ylab(expression(Rs[GPP] ~ (Pg ~ C ~ yr^{-1}))) +
  labs(x = element_blank(), y = element_blank()) +
  theme(legend.position = "none") -> rep_Rroot_Rsgpp_regression

ggdraw() +
  draw_plot(rep_froot_regression) +
  draw_plot(rep_Rroot_Rsgpp_regression, x = 0.45, y = 0.5, width = 0.5, height = 0.4) ->
  rep_froot_regression_draw



## plot rroot:rs representative
bind_rows(
  sub_srdb_fluxcom %>% select(FLUXCOM_GPP) %>% mutate(Source = "RrRs"),
  Fluxcon_gpp %>% select(FLUXCOM_GPP) %>% mutate(Source = "Global")) %>% 
  ggplot(aes(x = FLUXCOM_GPP, fill = factor(Source))) +
  theme(legend.position = c(0.75,0.75)) +
  # geom_violin(draw_quantiles = c(0.5))
  geom_density(alpha = 0.5) +
  scale_fill_manual(breaks = c("RrRs", "Global"),
                    labels = c(expression(R[root]~':'~R[S]), expression(Global)),
                    values = c("chocolate", "coral4"))+
  theme(
    legend.title = element_blank(),
    legend.text.align = 0)+
  labs(x=element_blank(), y=expression(Density)) ->
  rep_rroot_rs_gpp

## regresstion between ratios and GPP
lm(sub_srdb_fluxcom$RC_annual ~ sub_srdb_fluxcom$FLUXCOM_GPP) %>% summary()

sub_srdb_fluxcom %>% select(RC_annual, FLUXCOM_GPP) %>% 
  ggplot(aes(FLUXCOM_GPP, RC_annual)) +
  geom_point() +
  theme(
    legend.title = element_blank(),
    legend.text.align = 0)+
  labs(x=element_blank(), y=expression(R[root]~':'~R[S])) ->
  rep_Rroot_regression


## plot ra:GPP representative
bind_rows(
  sub_raGPP_fluxcom %>% select(FLUXCOM_GPP) %>% mutate(Source = "RaGPP"),
  Fluxcon_gpp %>% select(FLUXCOM_GPP) %>% mutate(Source = "Global")) %>% 
  ggplot(aes(x = FLUXCOM_GPP, fill = factor(Source))) +
  theme(legend.position = c(0.75,0.75)) +
  # geom_violin(draw_quantiles = c(0.5))
  geom_density(alpha = 0.5) +
  scale_fill_manual(breaks = c("RaGPP", "Global"),
                    labels = c(expression(R[A]~':'~GPP), expression(Global)),
                    values = c("chocolate", "coral4"))+
  theme(
    legend.title = element_blank(),
    legend.text.align = 0)+
  labs(x=expression(GPP~"("~g~C~m^{-2}~day^{-1}~")"), y=expression(Density)) ->
  rep_ra_gpp_gpp

## regresstion between ratios and GPP
lm(sub_raGPP_fluxcom$RaGPP_ratio ~ sub_raGPP_fluxcom$FLUXCOM_GPP) %>% summary()

sub_raGPP_fluxcom %>% select(RaGPP_ratio, FLUXCOM_GPP) %>% 
  ggplot(aes(FLUXCOM_GPP, RaGPP_ratio)) +
  geom_point() +
  # geom_smooth(method = "lm") +
  theme(
    legend.title = element_blank(),
    legend.text.align = 0)+
  labs(x=expression(GPP~"("~g~C~m^{-2}~day^{-1}~")"), y=expression(R[A]~':'~GPP)) ->
  rep_RaGPP_regression  


(rep_rroot_rs_gpp | rep_Rroot_regression) / (rep_rroot_ra_gpp | rep_froot_regression) / (rep_ra_gpp_gpp | rep_RaGPP_regression) +
  plot_annotation(tag_levels = 'a')

# ggsave("outputs/FigureSX.png", width = 9, height = 8)

```


```{r GPP by group plot}
# plot gpp by different group category
GPP %>% ggplot(aes(x = Group2, y = GPP, fill = as.factor(Group2))) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_jitter(shape = 16, position = position_jitter(0.2), col = "gray") +
  # geom_boxplot(width = 0.1) +
  theme(legend.title = element_blank()) +
  scale_fill_brewer(palette="Spectral") +
  xlab("GPP by different groups") +
  ylab(expression(GPP ~ "(Pg C" ~ yr^{-1} ~ ")")) -> gpp_by_group
```


```{r gpp group and its affect}
## gpp_results is from "Topdown_GlobalRs.Rmd"
## plot and show how aggregating through group could affect gpp bootdtrap output

bind_rows(
  gpp_results %>% select(`value` = GPP2) %>% 
  mutate(GPP_type = "GPP_Rs"),
  bootstrap_data2 %>%
  select(`GPP` = GPP_raw, `GPP[group]` = GPP_raw_group) %>%
  gather(GPP_type)) %>% 
  ggplot(aes(value, fill = GPP_type)) +
  geom_density(stat = "density", alpha = 0.65) +
  # theme(legend.position = c(0.70, 0.50)) +
  coord_cartesian(xlim = c(80, 210)) +
  scale_fill_manual("",
    labels = c(expression(GPP[lit]), expression(GPP[lit-group]), expression(GPP[Rs])),
    values=c("#E41A1C", "darkred", "skyblue")
  ) +
  theme(
    legend.title = element_blank(),
    legend.text.align = 0
  ) +
  ylab("Density") +
  xlab(expression(GPP ~ (Pg ~ C ~ yr^{-1}))) -> GPP_lit_group_plot
```

```{r outputs gpp group affect result}
gpp_by_group / GPP_lit_group_plot +
  plot_annotation(tag_levels = 'a')
# ggsave("outputs/FigureSX5.png", width = 8, height = 6)
```

Presumably, as a matter of mathematics, the ratio of global Rs to global GPP should be the linear combination (sum) of all terrestrial points Neither SRDB nor FLUXNET is an unbiased sample of the land surface, but it may be instructive to look at the Rs:GPP for three data sources:
* Global GPP estimates, comparing them against the mean Rs estimate (~85 Pg C);
* the SRDB itself, for sites at which both GPP and Rs were measured;
* and FLUXNET2015 Tier 1, using Rs data measured very close to the towers.

```{r, site-ratios, results='show'}
topdown <- expand.grid(GPP = GPP$GPP, Rs = GlobalRs$Rs)
topdown$Source <- paste0("Rs and GPP\nfrom the literature\n(N = ", nrow(topdown), ")")
topdown$Type <- "Global"
topdown$Scale <- "Global scale"
topdown_mean <- mean(topdown$Rs / topdown$GPP)

srdb_v4 %>%
  filter(!is.na(GPP), !is.na(Rs_annual), Ecosystem_state != "Managed") %>%
  select(Rs = Rs_annual, GPP, Biome, Ecosystem_type) ->
srdb
srdb$Source <- paste0("SRDB\n(N = ", nrow(srdb), ")")
srdb$Type <- "Observations"
srdb$Scale <- "Ecosystem scale"

read.csv("data/site-data-temp/srdb_filtered.csv", stringsAsFactors = FALSE) %>%
  select(Biome, Ecosystem_type, FLUXNET_SITE_ID, GPP = gpp_fluxnet, Rs = Rs_annual) %>%
  filter(!is.na(GPP), !is.na(Rs)) ->
fluxnet
fluxnet$Source <- paste0("FLUXNET\n(N = ", nrow(fluxnet), ")")
fluxnet$Type <- "Observations"
fluxnet$Scale <- "Ecosystem scale"

cmip6_global_data_file <- "cmip/CMIP6_annual_RstoGPP.csv"
if (file.exists(cmip6_global_data_file)) {
  read.csv(cmip6_global_data_file, stringsAsFactors = FALSE) %>%
    as_tibble() %>%
    filter(year >= 2005) %>%
    group_by(model, experiment, ensemble) %>%
    summarise(GPP = mean(GPP), Rs = mean(Rs)) %>%
    ungroup() ->
  model_global_ratios

  model_global_ratios %>%
    mutate(
      Scale = "Global scale",
      Source = paste0("CMIP6\n(N = ", nrow(model_global_ratios), ")"),
      Type = "Models"
    ) ->
  cmip_global
} else {
  cmip_global <- data.frame(
    Source = "CMIP6\n(N = 0)", Scale = "Global",
    Type = "Models", stringsAsFactors = FALSE
  )
}

cmip6_local_data_file <- "cmip/CMIP6_LatLon.csv"
if (file.exists(cmip6_local_data_file)) {
  read.csv(cmip6_local_data_file, stringsAsFactors = FALSE) %>%
    as_tibble() %>% 
     group_by(model, experiment, ensemble, Longitude, Latitude) %>%
    summarise(GPP = mean(GPP), Rs = mean(Rs)) %>%
    ungroup() -> 
  cmip6_local_data

  cmip6_local_data %>%
    select(experiment, model, ensemble) %>%
    distinct() %>%
    nrow() ->
  samples

  cmip6_local_data %>%
    mutate(
      Scale = "Ecosystem scale",
      Source = paste0("CMIP6\n(N = ", samples, ")"),
      Type = "Models"
    ) ->
  cmip_local
} else {
  cmip_local <- data.frame(
    Source = "CMIP6\n(N = 0)",
    Scale = "Ecosystem scale",
    Type = "Models",
    stringsAsFactors = FALSE
  )
}

bind_rows(topdown, srdb, fluxnet, cmip_global, cmip_local) %>%
  mutate(
    Ratio = Rs / GPP,
    Source = factor(Source, unique(Source)[c(2, 3, 5, 4, 1)])
  ) ->
ratios

ratios %>%
  group_by(Source) %>%
  summarise(
    Ratio_median = round(median(Ratio, na.rm = TRUE), 3),
    Ratio_sd = round(sd(Ratio, na.rm = TRUE), 3)
  ) 

p_ratios <- ggplot(ratios, aes(Source, Ratio, color = Type)) +
  geom_boxplot(show.legend = FALSE, col = "black", fill = "white", alpha = 1) +
  geom_quasirandom(alpha = I(0.20)) +
  scale_color_brewer(palette="Set1") +
  geom_hline(yintercept = topdown_mean, linetype = 2, color = "red", alpha = 0.33) +
  ylab(expression(Ratio ~ of ~ R[S] ~ to ~ GPP)) +
  xlab("") +
  coord_cartesian(ylim = c(0, 2)) +
  facet_grid(~Scale, scales = "free_x") +
  theme(strip.background = element_rect(fill = "white", linetype = "blank")) +
  guides(colour = guide_legend(override.aes = list(alpha = 1)))
print(p_ratios)

# ggsave("outputs/Figure2-1.png")

# Test whether global group is significantly higher than others
ratios %>%
  select(Source, Ratio, Scale, Type) %>%
  as_tibble() %>%
  na.omit() %>%
  mutate(globalcase = Type == "Global") ->
ratiotest
cat("Test whether global RS:GPP higher than all others:")
wilcox.test(Ratio ~ globalcase, data = ratiotest, alternative = "less")
cat("Test whether global RS:GPP higher than observations:")
ratiotest_obs <- filter(ratiotest, Type %in% c("Global", "Observations"))
wilcox.test(Ratio ~ globalcase, data = ratiotest_obs, alternative = "less")
```

# bootstrapping and show the representation affects

```{r bootstrap-creation}
# Bootstrap settings
RESAMPLE <- method3
RESAMPLE_noSD <- method1
N_SAMPLES <- 10000
sample_size <- 10
QUANTILES <- c(0.025, 0.5, 0.975) # Generate mean and 95% CI
set.seed(20190827)

# Rs and GPP draws used by bootstrapping GPP and Rs blocks below
bootstrap_data3 <- tibble(
  Rs_raw = sapply(rep(sample_size, N_SAMPLES), RESAMPLE, GlobalRs$Rs, GlobalRs$SD),
  Rs_raw_agg = sapply(rep(sample_size, N_SAMPLES), RESAMPLE, GlobalRs_agg$Rs, GlobalRs_agg$SD),
  GPP_raw = sapply(rep(sample_size, N_SAMPLES), RESAMPLE, GPP$GPP, GPP$SD),
  GPP_raw_agg_group = sapply(rep(sample_size, N_SAMPLES), RESAMPLE, GPP_agg_group$GPP, GPP_agg_group$SD),
)

sd(bootstrap_data3$Rs_raw_agg) # why is sd so small? not as expected
sd(bootstrap_data3$GPP_raw_agg_group) # why is sd so small? not as expected
```


## Bootstrapping GPP
```{r gpp-bootstrap}
# Bottom-up approach to estimate GPP
# Rroot = Rs * RC (Rroot to Rs ratio, from srdb)
# Froot = Rroot / Ra (root autotrophic respiration to total autotrophic respiration fraction/ratio, calculation please see plot_Rroot_Ra_ratio function)
# Fshoot = 1-Froot
# FsFr = (1-Froot)/Froot (FsFr = Rshoot/Rroot ratio)
# Ra = Rroot + Rshoot
# GPP = NPP + Rroot + Rshoot

# Step 1. Prepare samples of all variables that go into the GPP calculation
bootstrap_data3 %>%
  mutate(
    NPP = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, NPP$NPP),

    # Rroot-to Rs ratio (RC)
    Rc_agr = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_agr_rc$RC_annual),

    # forests are separated into 3 groups
    Rc_df = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_df_rc$RC_annual),
    Rc_ef = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_ef_rc$RC_annual),
    Rc_mf = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_mf_rc$RC_annual),
    # other ecosystems
    Rc_gra = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_gra_rc$RC_annual),
    Rc_shr = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_shr_rc$RC_annual),
    # for other ecosystems we sample from all RC records
    Rc_rest = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_rest_rc$RC_annual),

    # Rroot-to-Ra ratio (Froot)
    Froot_df = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_df_froot),
    Froot_ef = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_ef_froot),
    Froot_mf = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_mf_froot),
    # for other ecosystems we sample from all RC records
    Froot_rest = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_rest_froot),

    Rc = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_RC$RC_annual),

    # Scenario 1: Froot not separated into different ecosystems
    Froot = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, Froot$Froot)
  ) -> bootstrap_gpp_draws

# update the function to get Rs_gpp based on 
calc_bootstrap_gpp_agg <- function(x) {
  x %>%
    mutate(
      # Scenario 2: Rc separate into groups: agriculture, deciduous forest, grassland, etc
      # all other ecosystems with obs<40, take samples from all RC records
      # area of rest ecosystem = 1-sum(agriculture, forest, grassland, etc)
      Rc2 = (Rc_agr * ag_area + Rc_df * df_area + Rc_ef * ef_area + Rc_mf * mf_area +
               Rc_gra * gra_area + Rc_shr * shr_area + Rc_rest *
               (1 - ag_area - df_area - ef_area - mf_area - gra_area - shr_area)),
      
      FsFr = (1 - Froot) / Froot, # calculate Rshoot to Rroot ratio of scenario 1
      # Bottom up estimate of GPP, scenario 1
      Rroot = Rs_raw * Rc, # root respiration
      Rshoot = Rroot * FsFr, # shoot respiration
      
      # Scenario 2: separate into different ecosystems and weighted by area
      # Froot %>% count(Ecosystem)
      Froot2 = Froot_df * df_area + Froot_ef * ef_area + Froot_mf * mf_area +
        Froot_rest * (1 - df_area - ef_area - mf_area),
      FsFr2 = (1 - Froot2) / Froot2, # Rshoot to Rroot ratio of scenario 2
      # Bottom up estimate of GPP, scenario 2
      Rroot2 = Rs_raw * Rc2,
      Rshoot2 = Rroot2 * FsFr2,
      
      # Bottom up estimate of GPP, scenario 4 - using aggregated Rs
      Rroot4 = Rs_raw_agg * Rc2,
      Rshoot4 = Rroot4 * FsFr2,
      
      GPP = NPP + Rroot + Rshoot, # scenario 1: GPP
      GPP2 = NPP + Rroot2 + Rshoot2, # scenario 2: GPP2
      GPP4 = NPP + Rroot4 + Rshoot4 # scenario 4: GPP4
    )
}
# Now we can run the initial GPP bootstrapping
gpp_results_agg <- calc_bootstrap_gpp_agg(bootstrap_gpp_draws)
```

## boottrapping Rs
```{r rs-bootstrap}
# topdown approach to estimate Rs
# Ra = GPP * Ra_GPP_ratio (calculated based on data from a
#   summary paper and from srdb; see plot_RaGPP function)
# Ra = GPP - NPP (two approaches yield very similar results;
#   we thus used the average of two estimates)

# Step 1. Prepare samples of all variables that go into the Rs calculation
bootstrap_data3 %>%
  mutate(
    # NPP and NPP consumed by fire etc.
    NPP = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, NPP$NPP),

    # Carbon comsumed by herbivores (mean:2.2, SD: 0.2), from
    # Doughty et al., Agricultural net primary production in relation to that liberated by the extinction of Pleistocene mega-herbivores:
    # an estimate of agricultural carrying capacity? ERL 5, 044001 (2010).
    # https://iopscience.iop.org/article/10.1088/1748-9326/5/4/044001/meta
    HerbComsum = rnorm(n = N_SAMPLES, mean = 2.2, sd = 0.2),
    # Fire = sample(var_Fire$Amount_Pg, N_SAMPLES, replace = TRUE),
    Fire = sapply(rep(sample_size, N_SAMPLES), RESAMPLE, var_Fire$Amount_Pg, var_Fire$SD),
    # Carbon going to terrestrial as carbon sink (mean = 2.1, SD = 0.28), from
    # Le Quéré, C. et al. Global carbon budget 2015. Earth Syst. Sci. Data 7, 349–396 (2015).
    sink = rnorm(n = N_SAMPLES, mean = 2.10, sd = 0.28),
    # Dissolved organic carbon exports (Table S3)
    DOC = rnorm(n = N_SAMPLES, mean = 2.3, sd = 0.25),
    # Carbon of BVOCs (there is no SD information, take SD = Mean*0.1, see Table S3)
    BVOCs = rnorm(n = N_SAMPLES, mean = 1.007, sd = 0.1),

    # Step 2. Ra-to-GPP ratio boosting
    # RaGpp = sample(var_RaGpp$RaGPP_ratio, N_SAMPLES, replace = TRUE),
    RaGpp_df = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_RaGpp_df),
    RaGpp_ef = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_RaGpp_ef),
    RaGpp_mf = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_RaGpp_mf),
    RaGpp_gra = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_RaGpp_gra),
    # for other ecosystems, we use samples from all Froot
    RaGpp_rest = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_RaGpp$RaGPP_ratio),

    # Step 3. Sampling Rroot-to-Ra ratio (Froot)
    Froot_df = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_df_froot),
    Froot_ef = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_ef_froot),
    Froot_mf = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_mf_froot),
    Froot_gra = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_gra_froot),
    # for other ecosystems, we use samples from all Froot
    Froot_rest = sapply(rep(sample_size, N_SAMPLES), RESAMPLE_noSD, var_rest_froot)
  ) ->
bootstrap_rs_draws

## update the bootstrap function for scenario 4
# Do the static (no random draws) steps that calculate Rs
# We define a function to re-use later for variance decomposition
calc_bootstrap_rs_agg <- function(x) {
  x %>%
    mutate(
      # Scenario 1
      RaGpp = RaGpp_rest,
      # Scenario 2
      RaGpp2 = RaGpp_df * df_area + RaGpp_ef * ef_area + RaGpp_mf * mf_area +
        RaGpp_gra * gra_area + RaGpp_rest * (1 - df_area - ef_area - mf_area - gra_area),
      
      # get Froot (weight by area of ecosystem type)
      Froot2 = Froot_df * df_area + Froot_ef * ef_area + Froot_mf * mf_area +
        Froot_rest * (1 - df_area - ef_area - mf_area),
      
      Ra1 = GPP_raw * RaGpp, # first way to calculate Ra, Ra = GPP * RaGPP
      Ra2 = GPP_raw - NPP, # second way to calculate Ra, Ra = GPP - NPP
      Ra3 = if_else(Ra2 < 0, Ra1, Ra2),
      Ra_avg1 = (Ra1 + Ra3) / 2,
      
      Rroot = Ra_avg1 * Froot2,
      Rshoot = Ra_avg1 * (1 - Froot2),
      Rs_topdown = NPP - HerbComsum - Fire - sink - DOC - BVOCs + Rroot,
      
      # Scenario 2
      Ra4 = GPP_raw * RaGpp2, # first way to calculate Ra, Ra = GPP * RaGPP
      Ra_avg2 = (Ra4 + Ra3) / 2,
      Rroot2 = Ra_avg2 * Froot2,
      Rshoot2 = Ra_avg2 * (1 - Froot2),
      Rs_topdown2 = NPP - HerbComsum - Fire - sink - DOC - BVOCs + Rroot2,
      
      
      
      # Scenario 4
      Ra5 = GPP_raw_agg_group * RaGpp, # first way to calculate Ra, Ra = GPP * RaGPP
      Ra6 = GPP_raw_agg_group - NPP, # second way to calculate Ra, Ra = GPP - NPP
      Ra7 = if_else(Ra6 < 0, Ra5, Ra6),
      
      Ra8 = GPP_raw_agg_group * RaGpp2, # first way to calculate Ra, Ra = GPP * RaGPP
      Ra_avg4 = (Ra8 + Ra7) / 2,
      Rroot4 = Ra_avg4 * Froot2,
      Rs_topdown4 = NPP - HerbComsum - Fire - sink - DOC - BVOCs + Rroot4
    )
}

rs_results_agg <- calc_bootstrap_rs_agg(bootstrap_rs_draws)
```


```{r plot and show the representative final results}
# plot 1
gpp_results_agg %>%
  select(`GPP[lit]` = GPP_raw, `GPP[Rs]` = GPP2) %>%
  gather(GPP_type) %>% 
  ggplot(aes(value, fill = GPP_type)) +
  geom_density(stat = "density", alpha = 0.65) +
  coord_cartesian(xlim = c(80, 210)) +
  scale_fill_manual("",
    labels = c(expression(GPP[lit]), expression(GPP[Rs])),
    values=c("#E41A1C", "skyblue")
  ) +
  theme(
    legend.position = c(0.80, 0.50),
    legend.title = element_blank(),
    legend.text.align = 0
  ) +
  ylab("Density") +
  xlab(expression(GPP ~ (Pg ~ C ~ yr^{-1}))) -> GPP_plot1
# print(GPP_plot1)

# plot 2
gpp_results_agg %>%
  select(`GPP[agg]` = GPP_raw_agg_group, `GPP[Rs-agg]` = GPP4) %>%
  gather(GPP_type) %>% 
  ggplot(aes(value, fill = GPP_type)) +
  geom_density(stat = "density", alpha = 0.65) +
  coord_cartesian(xlim = c(80, 210)) +
  scale_fill_manual("",
    labels = c(expression(GPP[lit-agg]), expression(GPP[Rs-agg])),
    values=c("#E41A1C", "skyblue")
  ) +
  theme(
    legend.position = c(0.80, 0.50),
    legend.title = element_blank(),
    legend.text.align = 0
  ) +
  ylab("Density") +
  xlab(expression(GPP ~ (Pg ~ C ~ yr^{-1}))) -> GPP_plot2

# plot3
rs_results_agg %>%
  select(`Rs[lit]` = Rs_raw, `Rs[GPP]` = Rs_topdown2) %>%
  gather(Rs_type) %>% 
  ggplot(aes(value, fill = Rs_type)) +
  geom_density(stat = "density", alpha = 0.65) +
  coord_cartesian(xlim = c(50, 110)) +
  scale_fill_manual("",
    labels = c(expression(Rs[GPP]), expression(Rs[lit])),
    values=c("#E41A1C", "skyblue")
  ) +
  theme(
    legend.position = c(0.85, 0.50),
    legend.title = element_blank(),
    legend.text.align = 0
  ) +
  ylab("Density") +
  xlab(expression(Rs ~ (Pg ~ C ~ yr^{-1}))) -> Rs_plot3


# plot4
rs_results_agg %>%
  select(`Rs[lit]` = Rs_raw_agg, `Rs[GPP]` = Rs_topdown4) %>%
  gather(Rs_type) %>% 
  ggplot(aes(value, fill = Rs_type)) +
  geom_density(stat = "density", alpha = 0.65) +
  coord_cartesian(xlim = c(50, 110)) +
  scale_fill_manual("",
    labels = c(expression(Rs[GPP-agg]), expression(Rs[lit-agg])),
    values=c("#E41A1C", "skyblue")
  ) +
  theme(
    legend.position = c(0.80, 0.50),
    legend.title = element_blank(),
    legend.text.align = 0
  ) +
  ylab("Density") +
  xlab(expression(Rs ~ (Pg ~ C ~ yr^{-1}))) -> Rs_plot4


```


```{r, fig.height=6, fig.width=8}
(GPP_plot1 | Rs_plot3) / (GPP_plot2 | Rs_plot4) +
  plot_annotation(tag_levels = 'a')
# ggsave("outputs/FigureSXXX.png", width = 8, height = 6)
```


