---
title: "GlobalC"
author: "Jinshi Jian and co-authors"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
  word_document: default
---

```{r preliminaries, message=FALSE, echo=FALSE}
# Constants
OUTPUT_DIR <- "outputs/"
DATA_DIR <- "data"
# Create output and log folders if they do not exist
if (!file.exists(OUTPUT_DIR)) dir.create(OUTPUT_DIR)
library(kableExtra)
library(data.table)
library(lubridate)
library(knitr)
library(ggpubr)
library(ggplot2)
theme_set(theme_bw())
library(cowplot)
library(ggmap)
library(maps)
library(mapdata)
library(ggExtra)
library(tidyr)
library(hexbin)
library(dplyr)
library(ggbeeswarm)
library(ggridges)
# Source all needed functions
source("functions.R")
source("functions-stats.R")
source("functions-resampling.R")
source("functions-calc_bootstrap.R")
library(randomForest)
library(patchwork)

# Set chunks defaults; these options will be applied to all subsequent chunks
knitr::opts_chunk$set(
  results = "hide", message = TRUE, include = TRUE,
  echo = FALSE, warning = FALSE,
  fig.height = 4, fig.width = 8, cache = FALSE
)
```

## load data
```{r}
drake::readd(mgrsd_climate) -> mgrsd_climate #read mgrsd data from drake (linked with environmental factors in drake)
mgrsd_climate$EVI[mgrsd_climate$EVI < 0 & !is.na(mgrsd_climate$EVI)] <- 0 # for those EVI is null and < 0 (covered by snow etc), set to 0
drake::readd(global_land) -> global_land # read global 0.5 by 0.5 resolution data for upscaling RF model to predict global RS
```

## prepare mgrsd for modeling
```{r}
# Prepare mgrsd for RF modeling, need remove NA data, exclude Rs_annual < 0 value, and need to remove manupulation != "none"
mgrsd_climate %>% 
  dplyr::select(Rs_Norm, Meas_Month2, Manipulation, Study_number, 
                BM_aboveground, BM_belowground, N_dep_half_deg, bd_soilgrids, clay_soilgrids, soc_soilgrids, Pmonth, Tmonth, EVI) %>% 
  filter(Manipulation == "None" & Rs_Norm >= 0) %>% 
  # filter(!is.na(EVI))
  na.omit() ->
  sub_mgrsd
```

## testing code
### plot Rs_norm vs all those 9 variables used in the RF model
```{r}
plot_scatter <- function(){
  for(i in 4:ncol(sub_mgrsd)){
    subdata <- sub_mgrsd[,c(1,i)]
    var_factor <- colnames(subdata)[2]
    colnames(subdata) <- c("Rs", "Factor")
    lm_mod <- lm(Rs ~ Factor, data = subdata)
    R2 <- summary(lm_mod)$r.squared %>% round(5)
    subdata %>% 
      ggplot(aes(Factor, Rs)) +
      geom_point(alpha = 0.25) +
      ggtitle(paste0(var_factor, "(R2=", R2, ")")) ->
      p
    plot(p)
    }
  }

plot_scatter()
```

## random forest modeling
```{r}
set.seed(20201122)
## 75% of the sample size
## set the seed to make your partition reproducible
smp_size <- floor(0.80 * nrow(sub_mgrsd))
train_ind <- sample(seq_len(nrow(sub_mgrsd)), size = smp_size)

train <- sub_mgrsd[train_ind, ]
test <- sub_mgrsd[-train_ind, ]

# rf for train
rf <- randomForest(Rs_Norm ~ BM_aboveground + BM_belowground +
                     N_dep_half_deg + bd_soilgrids + clay_soilgrids + soc_soilgrids + EVI + Tmonth + Pmonth, 
                   data = train,
                   importance = TRUE, 
                   proximity = TRUE,
                   ntree = 200,
                   mtry = 4,
                   na.action = na.exclude)   

plot(rf)
importance(rf)
varImpPlot(rf)
```

```{r, fig.width=6, fig.height=8}
# plot train vs test
# Using RF model predict Rs_annual for train dataset and test dataset and evaluate the model performance
train %>% 
  mutate(prediction = predict(rf)) ->
  train

R2_train <- summary(lm(Rs_Norm ~ prediction, data = train))$r.squared %>% round(2)
train %>% 
  ggplot(aes(prediction, Rs_Norm)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm") +
  geom_abline(slope = 1, intercept = 0, col = "red", linetype = "dashed") +
  ggtitle(paste0("(a) train data (n = 21771, 80%)")) +
  annotate("text", x = 0, y = 27.5, label = paste0("italic(R) ^ 2"), parse = TRUE, col = "black", hjust = 0) +
  annotate("text", x = 0.75, y = 27.5, label = paste0("=", R2_train), col = "black", hjust = 0) +
  labs(x = expression(Predicted~R[S]~(g~C~m^{2}~day^{-1})),
       y = expression(Measured~R[S]~(g~C~m^{2}~day^{-1}))) ->
  train_plot

test %>% 
  mutate(prediction = predict(rf, test)) ->
  test

R2_test <- summary(lm(Rs_Norm ~ prediction, data = test))$r.squared %>% round(2)
test %>% 
  ggplot(aes(prediction, Rs_Norm)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm") +
  geom_abline(slope = 1, intercept = 0, col = "red", linetype = "dashed") +
  ggtitle(paste0("(b) test data (n = 5443, 20%)")) +
  annotate("text", x = 0, y = 27.5, label = paste0("italic(R) ^ 2"), parse = TRUE, col = "black", hjust = 0) +
  annotate("text", x = 0.6, y = 27.5, label = paste0("=", R2_test), col = "black", hjust = 0) +
  labs(x = expression(Predicted~R[S]~(g~C~m^{2}~day^{-1})),
       y = expression(Measured~R[S]~(g~C~m^{2}~day^{-1}))) ->
  test_plot

train_plot / test_plot
```

## test the model performance by each month
```{r}
# rf for each month
rf_month <- function(mon){
  mon_mgrsd <- sub_mgrsd %>% filter(Meas_Month2 == mon)
  randomForest(Rs_Norm ~ BM_aboveground + BM_belowground +
                     N_dep_half_deg + bd_soilgrids + clay_soilgrids + soc_soilgrids + EVI + Tmonth + Pmonth, 
                   data = mon_mgrsd,
                   importance = TRUE, 
                   proximity = TRUE,
                   ntree = 200,
                   mtry = 4,
                   na.action = na.exclude)  
  }
 
rf_Jan <- rf_month(1)
plot(rf_Jan)
importance(rf_Jan)
varImpPlot(rf_Jan)

rf_Feb <- rf_month(2)
plot(rf_Feb)
importance(rf_Feb)
varImpPlot(rf_Feb)

rf_Mar <- rf_month(3)
rf_Apr <- rf_month(4)
rf_May <- rf_month(5)
rf_Jun <- rf_month(6)
rf_Jul <- rf_month(7)
rf_Aug <- rf_month(8)
rf_Sep <- rf_month(9)
rf_Oct <- rf_month(10)
rf_Nov <- rf_month(11)
rf_Dec <- rf_month(12)

```

## Create a functiin to test RF model by month
```{r}
eval_mon <- function(mon) {
  mon_mgrsd <- sub_mgrsd %>% filter(Meas_Month2 == mon) %>% 
    mutate(prediction = if (mon == 1) {
               predict(rf_Jan)}
           else if(mon == 2){
             predict(rf_Feb)
           } else if (mon == 3) {
             predict(rf_Mar)
           } else if (mon == 4) {
             predict(rf_Apr)
           } else if (mon == 5) {
             predict(rf_May)
           } else if (mon == 6) {
             predict(rf_Jun)
           } else if (mon == 7) {
             predict(rf_Jul)
           } else if (mon == 8) {
             predict (rf_Aug)
           } else if (mon == 9) {
             predict (rf_Sep)
           } else if (mon == 10) {
             predict (rf_Oct)
           } else if (mon == 11) {
             predict (rf_Nov)
           } else if (mon == 12) {
             predict (rf_Dec)
           }
      ) ->
    mon_mgrsd
  
  R2_mon <- summary(lm(Rs_Norm ~ prediction, data = mon_mgrsd))$r.squared %>% round(3)
  mon_mgrsd %>% 
    ggplot(aes(prediction, Rs_Norm)) +
    geom_point(alpha = 0.25) +
    geom_smooth(method = "lm") +
    geom_abline(slope = 1, intercept = 0, col = "red", linetype = "dashed") +
    ggtitle(paste0("Month=", i, ", R^2", "=", R2_mon, ", n=", nrow(mon_mgrsd))) +
    labs(x = expression(Predicted~R[S]~(g~C~m^{2}~day^{-1})),
         y = expression(Measured~R[S]~(g~C~m^{2}~day^{-1}))) ->
    mon_plot
  print(mon_plot)
}

for(i in 1:12) {
  eval_mon(i)}
```

## QA/QC to check pottential issues (no problem was found)
```{r}
sub_mgrsd %>% filter(Meas_Month2 == 6) -> mon_mgrsd

sub_mgrsd$EVI %>% min()
sub_mgrsd$EVI %>% max()

mon_mgrsd %>% 
  mutate(prediction = predict(rf_Jun, mon_mgrsd)) ->
  mon_mgrsd

mon_mgrsd %>% filter(prediction > 10)

# Study number 8887
mgrsd_climate %>% 
  filter(Study_number == 8887) %>% 
  ggplot(aes(Tmonth, Tsoil)) +
  geom_point()

mgrsd_climate %>% 
  filter(Study_number == 8887) %>% 
  ggplot(aes(Tsoil, Rs_Norm)) +
  geom_point()

mgrsd_climate %>% 
  filter(Study_number == 8887) %>% 
  ggplot(aes(Tmonth, Rs_Norm)) +
  geom_point()
  
mgrsd_climate %>% 
  dplyr::select(Soil_BD, bd_soilgrids) %>% 
  na.omit() %>% 
  ggplot(aes(Soil_BD, bd_soilgrids)) +
  geom_point()

raster("data/EVIs/EVI_mean_month_1.tif") %>% 
  image()
```

## Prepare global 0.5 resolution data for predicting global Rs by month
```{r}
hist(global_land$EVI_1)
global_land %>% 
  ggplot(aes(Longitude, Latitude, col = EVI_1)) + 
  geom_point()

global_land %>% 
  ggplot(aes(Longitude, Latitude, col = prec1)) +
  geom_point()

global_land %>% 
  filter(is.na(tmean6)) %>% 
  ggplot(aes(Longitude, Latitude)) +
  geom_point()

global_land %>% 
  filter(!is.na(tmean6)) %>% 
  ggplot(aes(Longitude, Latitude)) +
  geom_point()
```

## Handle NA values
```{r}
# cells with no temperature records removed and save to global_sub
global_land %>% 
  filter(!is.na(tmean6)) ->
  global_sub

global_sub %>% 
  filter(is.na(tmean1) | is.na(tmean2) | is.na(tmean3)| is.na(tmean4) | is.na(tmean5)| is.na(tmean6)
         | is.na(tmean7)| is.na(tmean8)| is.na(tmean9)| is.na(tmean10)| is.na(tmean11))

global_sub %>% filter(is.na(tmean12))

# use the average of Nov and Jan to get Dec for those 301 NA values
global_sub$tmean12[is.na(global_sub$tmean12)] = (global_sub$tmean11[is.na(global_sub$tmean12)] + global_sub$tmean11[is.na(global_sub$tmean12)])/2

global_sub %>% 
  filter(is.na(prec1) | is.na(prec2) | is.na(prec3)| is.na(prec4) | is.na(prec5)| is.na(prec6)
         | is.na(prec7)| is.na(prec8)| is.na(prec9)| is.na(prec10)| is.na(prec11) | is.na(prec12))

## Take care of the EVI NA issue
global_sub$EVI_1[is.nan(global_sub$EVI_1) | global_sub$EVI_1 < 0] = 0
global_sub$EVI_2[is.nan(global_sub$EVI_2) | global_sub$EVI_2 < 0] = 0
global_sub$EVI_3[is.nan(global_sub$EVI_3) | global_sub$EVI_3 < 0] = 0
global_sub$EVI_4[is.nan(global_sub$EVI_4) | global_sub$EVI_4 < 0] = 0
global_sub$EVI_5[is.nan(global_sub$EVI_5) | global_sub$EVI_5 < 0] = 0
global_sub$EVI_6[is.nan(global_sub$EVI_6) | global_sub$EVI_6 < 0] = 0
global_sub$EVI_7[is.nan(global_sub$EVI_7) | global_sub$EVI_7 < 0] = 0
global_sub$EVI_8[is.nan(global_sub$EVI_8) | global_sub$EVI_8 < 0] = 0
global_sub$EVI_9[is.nan(global_sub$EVI_9) | global_sub$EVI_9 < 0] = 0
global_sub$EVI_10[is.nan(global_sub$EVI_10) | global_sub$EVI_10 < 0] = 0
global_sub$EVI_11[is.nan(global_sub$EVI_11) | global_sub$EVI_11 < 0] = 0
global_sub$EVI_12[is.nan(global_sub$EVI_12) | global_sub$EVI_12 < 0] = 0

## Soil data NA issue
global_sub %>% 
  filter(is.na(BM_aboveground_global)) 

global_sub %>% 
  filter(is.na(BM_belowground_global)) 

global_sub %>% 
  filter(is.na(N_dep_global)) 

global_sub %>% 
  filter(is.na(bd_soilgrids_global)) 

global_sub %>% 
  filter(is.na(clay_soilgrids_global)) 

global_sub %>% 
  filter(is.na(soc_soilgrids_global)) 

global_sub %>% 
  ggplot(aes(BM_belowground_global, soc_soilgrids_global)) +
  geom_point()

global_sub %>% 
  dplyr::select(BM_belowground_global, soc_soilgrids_global) %>% 
  filter(is.na(soc_soilgrids_global)) %>% 
  count(BM_belowground_global)

global_sub$bd_soilgrids_global %>% min(na.rm = TRUE)
global_sub$clay_soilgrids_global %>% min(na.rm = TRUE)
global_sub$soc_soilgrids_global %>% min(na.rm = TRUE)

global_sub %>% mutate(bd_soilgrids_global = bd_soilgrids_global / 1000) -> global_sub
global_sub$bd_soilgrids_global[is.na(global_sub$bd_soilgrids_global)] = 0
global_sub$clay_soilgrids_global[is.na(global_sub$clay_soilgrids_global)] = 0
global_sub$soc_soilgrids_global[is.na(global_sub$soc_soilgrids_global)] = 0

global_sub %>% 
  filter(is.na(bd_soilgrids_global)) %>% 
  ggplot(aes(Longitude, Latitude)) +
  geom_point()

global_sub %>% 
  na.omit() %>% 
  ggplot(aes(Longitude, Latitude)) +
  geom_point()
```

```{r}
# change column names of the global data
colnames(global_sub)[39:44] <- c("BM_aboveground", "BM_belowground", "N_dep_half_deg", "bd_soilgrids", "clay_soilgrids", "soc_soilgrids")
```

```{r}
# predict global Rs from Jan to Dec
global_sub %>% 
  mutate(EVI = EVI_1, Pmonth = prec1, Tmonth = tmean1) %>% 
  select(Latitude, Longitude, BM_aboveground, BM_belowground,
                     N_dep_half_deg, bd_soilgrids, clay_soilgrids, soc_soilgrids, EVI, Tmonth, Pmonth) ->
  global_sub1

global_sub %>% 
  mutate(EVI = EVI_2, Pmonth = prec2, Tmonth = tmean2) %>% 
  select(Latitude, Longitude, BM_aboveground, BM_belowground,
                     N_dep_half_deg, bd_soilgrids, clay_soilgrids, soc_soilgrids, EVI, Tmonth, Pmonth) ->
  global_sub2

global_sub %>% 
  mutate(EVI = EVI_3, Pmonth = prec3, Tmonth = tmean3) %>% 
  select(Latitude, Longitude, BM_aboveground, BM_belowground,
                     N_dep_half_deg, bd_soilgrids, clay_soilgrids, soc_soilgrids, EVI, Tmonth, Pmonth) ->
  global_sub3

global_sub %>% 
  mutate(EVI = EVI_4, Pmonth = prec4, Tmonth = tmean4) %>% 
  select(Latitude, Longitude, BM_aboveground, BM_belowground,
                     N_dep_half_deg, bd_soilgrids, clay_soilgrids, soc_soilgrids, EVI, Tmonth, Pmonth) ->
  global_sub4

global_sub %>% 
  mutate(EVI = EVI_5, Pmonth = prec5, Tmonth = tmean5) %>% 
  select(Latitude, Longitude, BM_aboveground, BM_belowground,
                     N_dep_half_deg, bd_soilgrids, clay_soilgrids, soc_soilgrids, EVI, Tmonth, Pmonth) ->
  global_sub5

global_sub %>% 
  mutate(EVI = EVI_6, Pmonth = prec6, Tmonth = tmean6) %>% 
  select(Latitude, Longitude, BM_aboveground, BM_belowground,
                     N_dep_half_deg, bd_soilgrids, clay_soilgrids, soc_soilgrids, EVI, Tmonth, Pmonth) ->
  global_sub6

global_sub %>% 
  mutate(EVI = EVI_7, Pmonth = prec7, Tmonth = tmean7) %>% 
  select(Latitude, Longitude, BM_aboveground, BM_belowground,
                     N_dep_half_deg, bd_soilgrids, clay_soilgrids, soc_soilgrids, EVI, Tmonth, Pmonth) ->
  global_sub7

global_sub %>% 
  mutate(EVI = EVI_8, Pmonth = prec8, Tmonth = tmean8) %>% 
  select(Latitude, Longitude, BM_aboveground, BM_belowground,
                     N_dep_half_deg, bd_soilgrids, clay_soilgrids, soc_soilgrids, EVI, Tmonth, Pmonth) ->
  global_sub8

global_sub %>% 
  mutate(EVI = EVI_9, Pmonth = prec9, Tmonth = tmean9) %>% 
  select(Latitude, Longitude, BM_aboveground, BM_belowground,
                     N_dep_half_deg, bd_soilgrids, clay_soilgrids, soc_soilgrids, EVI, Tmonth, Pmonth) ->
  global_sub9

global_sub %>% 
  mutate(EVI = EVI_10, Pmonth = prec10, Tmonth = tmean10) %>% 
  select(Latitude, Longitude, BM_aboveground, BM_belowground,
                     N_dep_half_deg, bd_soilgrids, clay_soilgrids, soc_soilgrids, EVI, Tmonth, Pmonth) ->
  global_sub10

global_sub %>% 
  mutate(EVI = EVI_11, Pmonth = prec11, Tmonth = tmean11) %>% 
  select(Latitude, Longitude, BM_aboveground, BM_belowground,
                     N_dep_half_deg, bd_soilgrids, clay_soilgrids, soc_soilgrids, EVI, Tmonth, Pmonth) ->
  global_sub11

global_sub %>% 
  mutate(EVI = EVI_12, Pmonth = prec12, Tmonth = tmean12) %>% 
  select(Latitude, Longitude, BM_aboveground, BM_belowground,
                     N_dep_half_deg, bd_soilgrids, clay_soilgrids, soc_soilgrids, EVI, Tmonth, Pmonth) ->
  global_sub12

```


```{r}
tibble(Rs_Jan = predict(rf_Jan, global_sub1),
       Rs_Feb = predict(rf_Feb, global_sub2),
       Rs_Mar = predict(rf_Mar, global_sub3),
       Rs_Apr = predict(rf_Apr, global_sub4),
       Rs_May = predict(rf_May, global_sub5),
       Rs_Jun = predict(rf_Jun, global_sub6),
       Rs_Jul = predict(rf_Jul, global_sub7),
       Rs_Aug = predict(rf_Aug, global_sub8),
       Rs_Sep = predict(rf_Sep, global_sub9),
       Rs_Oct = predict(rf_Oct, global_sub10),
       Rs_Nov = predict(rf_Nov, global_sub11),
       Rs_Dec = predict(rf_Dec, global_sub12)
       ) -> Global_Rs_prediction
```

```{r}
# Global Rs prediction standard diviation
tibble(
  Rs_sd_Jan = apply(predict(rf_Jan, global_sub1, predict.all=TRUE)$individual, 1, sd),
  Rs_sd_Feb = apply(predict(rf_Feb, global_sub2, predict.all=TRUE)$individual, 1, sd),
  Rs_sd_Mar = apply(predict(rf_Mar, global_sub3, predict.all=TRUE)$individual, 1, sd),
  Rs_sd_Apr = apply(predict(rf_Apr, global_sub4, predict.all=TRUE)$individual, 1, sd),
  Rs_sd_May = apply(predict(rf_May, global_sub5, predict.all=TRUE)$individual, 1, sd),
  Rs_sd_Jun = apply(predict(rf_Jun, global_sub6, predict.all=TRUE)$individual, 1, sd),
  Rs_sd_Jul = apply(predict(rf_Jul, global_sub7, predict.all=TRUE)$individual, 1, sd),
  Rs_sd_Aug = apply(predict(rf_Aug, global_sub8, predict.all=TRUE)$individual, 1, sd),
  Rs_sd_Sep = apply(predict(rf_Sep, global_sub9, predict.all=TRUE)$individual, 1, sd),
  Rs_sd_Oct = apply(predict(rf_Oct, global_sub10, predict.all=TRUE)$individual, 1, sd),
  Rs_sd_Nov = apply(predict(rf_Nov, global_sub11, predict.all=TRUE)$individual, 1, sd),
  Rs_sd_Dec = apply(predict(rf_Dec, global_sub12, predict.all=TRUE)$individual, 1, sd)
) -> Global_Rs_sd
 
```


```{r}
Global_Rs_prediction$Rs_mean <- apply(Global_Rs_prediction, 1, mean)
Global_Rs_prediction$Rs_sd_mean <- apply(Global_Rs_sd, 1, mean)
cbind(global_sub, Global_Rs_prediction) -> Global_Rs_prediction
```


```{r}
hist(Global_Rs_prediction$Rs_mean)
mean(Global_Rs_prediction$Rs_mean)
sd(Global_Rs_prediction$Rs_mean)
hist(Global_Rs_prediction$Rs_sd_mean)
mean(Global_Rs_prediction$Rs_sd_mean)
```

## Model evaluation: compare measured Rs vs predicted Rs from Jan to Dec
```{r}
# Winter
bind_rows(
  Global_Rs_prediction %>% 
    mutate(Type = "Global", Rs_mon = Rs_Dec) %>% 
    dplyr::select(Rs_mon, Type),
  
  sub_mgrsd %>% 
    mutate(Type = "MGRsD", Rs_mon = Rs_Norm) %>% 
    filter(Meas_Month2 == 12) %>% 
    dplyr::select(Rs_mon, Type) ) %>% 
  ggplot(aes(Rs_mon, Type, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
  xlim(0, 15) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Probability", direction = -1) +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(axis.text.y = element_text(angle = 0),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "none") ->
  Dec_comp

bind_rows(
  Global_Rs_prediction %>% 
    mutate(Type = "Global", Rs_mon = Rs_Jan) %>% 
    dplyr::select(Rs_mon, Type),
  
  sub_mgrsd %>% 
    mutate(Type = "MGRsD", Rs_mon = Rs_Norm) %>% 
    filter(Meas_Month2 == 1) %>% 
    dplyr::select(Rs_mon, Type) ) %>% 
  ggplot(aes(Rs_mon, Type, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
  xlim(0, 15) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Probability", direction = -1) +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "none") ->
  Jan_comp

bind_rows(
  Global_Rs_prediction %>% 
    mutate(Type = "Global", Rs_mon = Rs_Feb) %>% 
    dplyr::select(Rs_mon, Type),
  
  sub_mgrsd %>% 
    mutate(Type = "MGRsD", Rs_mon = Rs_Norm) %>% 
    filter(Meas_Month2 == 2) %>% 
    dplyr::select(Rs_mon, Type) ) %>% 
  ggplot(aes(Rs_mon, Type, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
  xlim(0, 15) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Probability", direction = -1) +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank()) ->
  Feb_comp

# Spring
bind_rows(
  Global_Rs_prediction %>% 
    mutate(Type = "Global", Rs_mon = Rs_Mar) %>% 
    dplyr::select(Rs_mon, Type),
  
  sub_mgrsd %>% 
    mutate(Type = "MGRsD", Rs_mon = Rs_Norm) %>% 
    filter(Meas_Month2 == 3) %>% 
    dplyr::select(Rs_mon, Type) ) %>% 
  ggplot(aes(Rs_mon, Type, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
  xlim(0, 15) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Probability", direction = -1) +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(axis.text.y = element_text(angle = 0),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "none") ->
  Mar_comp

bind_rows(
  Global_Rs_prediction %>% 
    mutate(Type = "Global", Rs_mon = Rs_Apr) %>% 
    dplyr::select(Rs_mon, Type),
  
  sub_mgrsd %>% 
    mutate(Type = "MGRsD", Rs_mon = Rs_Norm) %>% 
    filter(Meas_Month2 == 4) %>% 
    dplyr::select(Rs_mon, Type) ) %>% 
  ggplot(aes(Rs_mon, Type, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
  xlim(0, 15) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Probability", direction = -1) +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "none") ->
  Apr_comp

bind_rows(
  Global_Rs_prediction %>% 
    mutate(Type = "Global", Rs_mon = Rs_May) %>% 
    dplyr::select(Rs_mon, Type),
  
  sub_mgrsd %>% 
    mutate(Type = "MGRsD", Rs_mon = Rs_Norm) %>% 
    filter(Meas_Month2 == 5) %>% 
    dplyr::select(Rs_mon, Type) ) %>% 
  ggplot(aes(Rs_mon, Type, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
  xlim(0, 15) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Probability", direction = -1) +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank()) ->
  May_comp

# Summer
bind_rows(
  Global_Rs_prediction %>% 
    mutate(Type = "Global", Rs_mon = Rs_Jun) %>% 
    dplyr::select(Rs_mon, Type),
  
  sub_mgrsd %>% 
    mutate(Type = "MGRsD", Rs_mon = Rs_Norm) %>% 
    filter(Meas_Month2 == 6) %>% 
    dplyr::select(Rs_mon, Type) ) %>% 
  ggplot(aes(Rs_mon, Type, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
  xlim(0, 15) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Probability", direction = -1) +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(axis.text.y = element_text(angle = 0),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "none") ->
  Jun_comp

bind_rows(
  Global_Rs_prediction %>% 
    mutate(Type = "Global", Rs_mon = Rs_Jul) %>% 
    dplyr::select(Rs_mon, Type),
  
  sub_mgrsd %>% 
    mutate(Type = "MGRsD", Rs_mon = Rs_Norm) %>% 
    filter(Meas_Month2 == 7) %>% 
    dplyr::select(Rs_mon, Type) ) %>% 
  ggplot(aes(Rs_mon, Type, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
  xlim(0, 15) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Probability", direction = -1) +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "none") ->
  Jul_comp

bind_rows(
  Global_Rs_prediction %>% 
    mutate(Type = "Global", Rs_mon = Rs_Aug) %>% 
    dplyr::select(Rs_mon, Type),
  
  sub_mgrsd %>% 
    mutate(Type = "MGRsD", Rs_mon = Rs_Norm) %>% 
    filter(Meas_Month2 == 8) %>% 
    dplyr::select(Rs_mon, Type) ) %>% 
  ggplot(aes(Rs_mon, Type, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
  xlim(0, 15) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Probability", direction = -1) +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank()) ->
  Aug_comp

# Autumn
bind_rows(
  Global_Rs_prediction %>% 
    mutate(Type = "Global", Rs_mon = Rs_Sep) %>% 
    dplyr::select(Rs_mon, Type),
  
  sub_mgrsd %>% 
    mutate(Type = "MGRsD", Rs_mon = Rs_Norm) %>% 
    filter(Meas_Month2 == 9) %>% 
    dplyr::select(Rs_mon, Type) ) %>% 
  ggplot(aes(Rs_mon, Type, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
  xlim(0, 15) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Probability", direction = -1) +
  xlab(expression(R[S]~(g~C~m^{-2}~day^{-1}))) +
  ylab(element_blank()) +
  theme(axis.text.y = element_text(angle = 0),
        legend.position = "none") ->
  Sep_comp

bind_rows(
  Global_Rs_prediction %>% 
    mutate(Type = "Global", Rs_mon = Rs_Oct) %>% 
    dplyr::select(Rs_mon, Type),
  
  sub_mgrsd %>% 
    mutate(Type = "MGRsD", Rs_mon = Rs_Norm) %>% 
    filter(Meas_Month2 == 10) %>% 
    dplyr::select(Rs_mon, Type) ) %>% 
  ggplot(aes(Rs_mon, Type, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
  xlim(0, 15) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Probability", direction = -1) +
  xlab(expression(R[S]~(g~C~m^{-2}~day^{-1}))) +
  ylab(element_blank()) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "none") ->
  Oct_comp

bind_rows(
  Global_Rs_prediction %>% 
    mutate(Type = "Global", Rs_mon = Rs_Nov) %>% 
    dplyr::select(Rs_mon, Type),
  
  sub_mgrsd %>% 
    mutate(Type = "MGRsD", Rs_mon = Rs_Norm) %>% 
    filter(Meas_Month2 == 11) %>% 
    dplyr::select(Rs_mon, Type) ) %>% 
  ggplot(aes(Rs_mon, Type, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
  xlim(0, 15) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Probability", direction = -1) +
  xlab(expression(R[S]~(g~C~m^{-2}~day^{-1}))) +
  ylab(element_blank()) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) ->
  Nov_comp
```


```{r, fig.height=7, fig.width=8}
(Dec_comp | Jan_comp | Feb_comp) / (Mar_comp | Apr_comp | May_comp) /
  (Jun_comp | Jul_comp | Aug_comp) / (Sep_comp | Oct_comp | Nov_comp)
```

```{r}
# write global Rs production out
# write.csv(Global_Rs_prediction, "outputs/Global_Rs_prediction.csv", row.names = FALSE)
# write.table(Global_Rs_prediction, "outputs/Global_Rs_prediction.txt",sep = ",", row.names = FALSE)
```

