---
title: "Resampling_test_bbl"
output:
  word_document: default
  html_document: default
---

```{r preliminaries, message=FALSE, echo=FALSE}
# Constants
DATA_DIR <- 'data'

library(ggplot2)
theme_set(theme_bw())
library(cowplot)
library(tidyr)
library(dplyr)

# Set chunks defaults; these options will be applied to all subsequent chunks
knitr::opts_chunk$set(results = 'hide', message = TRUE, include = TRUE, 
                      echo = FALSE, warning = FALSE,
                      fig.height = 4, fig.width = 8, cache = FALSE)
```


## load data
```{r load data}
# input data
read_file <- function(x) read.csv(file.path(DATA_DIR, x), comment.char = "#", stringsAsFactors = FALSE)

GPP <- read_file('GlobalGPP_Sum.csv') # GPP estimates from published paper
GPP %>% filter(Pub_year > 1980 & Flag != "Model") -> GPP

GlobalRs <- read_file('GlobalRs.csv') # estimates of global Rs
GlobalRs %>% filter(Pub_year > 1980) -> GlobalRs
# (Ra/GPP) Piao, Shilong, et al. "Forest annual carbon cost: A global‐scale analysis of autotrophic respiration." Ecology 91.3 (2010): 652-661.
```


## Reasmple RSG and GPP
* Method 1 - ignore SD and resample from `r nrow (GlobalRs)` or `r nrow (GPP)` with replacement
* Method 2 - using SD, but calculated based on mean and coefficient of variability
* Method 3 - using SD, but using the maximum if missing
* Method 4 - using SD only when available, otherwise 0 SD

```{r}
set.seed(20191111)
N_SAMPLES = 5000

method1 <- function(n, x) {       # ignore all errors
  mean(sample(x, n, replace = TRUE))
}
method2 <- function(n, x, s, f = median) {  # replace NA errors with median error
  cv <- s / x
  cv <- replace_na(cv, f(cv, na.rm = TRUE))
  s <- if_else(is.na(s), x * cv, s)
  mean(rnorm(n, mean = x, sd = s))
}
method3 <- function(n, x, s) {    # replace NA errors with max error
  method2(n, x, s, max)
}
method4 <- function(n, x, s) {    # replace NA errors with zero
  method2(n, x, replace_na(s, 0))
}

RESAMPLE <- method3

# x <- RESAMPLE(...)


sample_size <- 10

# resample global soil respiration (RSG)
resample_RSG <- tibble(RSG1_no_SD = sapply(rep(sample_size, N_SAMPLES), method1, GlobalRs$Rs),
                       RSG2_miss_SD_median = sapply(rep(sample_size, N_SAMPLES), method2, GlobalRs$Rs, GlobalRs$SD),
                       RSG3_miss_SD_max = sapply(rep(sample_size, N_SAMPLES), method3, GlobalRs$Rs, GlobalRs$SD),
                       RSG4_miss_SD_zero = sapply(rep(sample_size, N_SAMPLES), method4, GlobalRs$Rs, GlobalRs$SD)
                       )

# resample gross primary production (GPP)
resample_GPP <- tibble(GPP1_no_SD = sapply(rep(sample_size, N_SAMPLES), method1, GPP$GPP),
                       GPP2_miss_SD_median = sapply(rep(sample_size, N_SAMPLES), method2, GPP$GPP, GPP$SD),
                       GPP3_miss_SD_max = sapply(rep(sample_size, N_SAMPLES), method3, GPP$GPP, GPP$SD),
                       GPP4_miss_SD_zero = sapply(rep(sample_size, N_SAMPLES), method4, GPP$GPP, GPP$SD)
                       )
```


## Compare the results of different re-sampling methods (N = `r N_SAMPLES`)
```{r}
# comparing RSG resample results
resample_RSG %>%
  gather(Rs_resample_type) %>% 
  ggplot(aes(value, fill = Rs_resample_type)) +
  theme_cowplot() +
  geom_density(stat = 'density', alpha = 0.5) +
  #  theme(legend.position = c(0.75, 0.75)) +
  # reverse the color scale so that teal (color 1) is always observations, 
  # while pink is always implied fluxes
  scale_fill_discrete("", h.start = 180) + scale_color_discrete(h.start = 180) +
  ylab("Density") +
  xlab(expression(R[S]~(Pg~C~yr^{-1}))) 

# comparing GPP resample results
resample_GPP %>%
  gather(GPP_resample_type) %>% 
  ggplot(aes(value, fill = GPP_resample_type)) +
  theme_cowplot() +
  geom_density(stat = 'density', alpha = 0.5) +
  #  theme(legend.position = c(0.75, 0.75)) +
  # reverse the color scale so that teal (color 1) is always observations, 
  # while pink is always implied fluxes
  scale_fill_discrete("", h.start = 180) + scale_color_discrete(h.start = 180) +
  ylab("Density") +
  xlab(expression(GPP~(Pg~C~yr^{-1}))) 
```



## test effect of sample size
```{r, fig.height=6, fig.width=8}
set.seed(20191114)

sample_size_tst(3, 50000)
sample_size_tst(5, 50000)
sample_size_tst(10, 50000)
sample_size_tst(20, 50000)
```


## test effect of resample time
```{r, fig.height=6, fig.width=8}
set.seed(20191114)

sample_size_tst(10, 1000)
sample_size_tst(10, 5000)
sample_size_tst(10, 10000)
sample_size_tst(10, 50000)
```

