---
title: "TopdownApproach"
author: "Jinshi"
date: "May 19, 2019"
output:
  html_document: default
  word_document: default
---


## Abbreviation
* Rroot - root respiration
* Rshoot - shoot respiration
* Rs - soil respiration
* RSG - global annual soil respiration 
* Ra - autotrophic respiration (Rroot + Rshoot)
* GPP - gross primary production
* NPP - net primary production
* NEP - net ecosystem production

* Froot - Rroot / Ra (root autotrophic respiration to total autotrophic respiration fraction/ratio, calculation please see plot_froot function)
* Fshoot - Rshoot / Ra (1-Froot)
* FsFr - (1-Froot)/Froot (FsFr = Rshoot/Rroot ratio)
* RC - (root respiration / soil respiration)
* RaGPP - (Ra/GPP) ratio


## Assumptions
* __Bottom-up approach:__ 
  1. We collected published global soil respiration estimates (mostly is based on bottom up field Rs measurements)
  2. We then partition Rs into root respiration and belowground heterotrophic resipiration
  3. We also collected published NPP estimates in the past decades from ITO et al. (2011)
  4. We then collected (and calculated) Rroot/Ra ratio, according to the root respiration, GPP can be calculates as (NPP + Rroot + Rshoot)
  5. We then compare this bottomup GPP with published GPP in the past decades (n=65) to see how much probability they will match

* __Top-down approach:__
  1. We first collect global GPP estimates in the past decades (most is top-dow remote sensing estimates)
  2. We then estimate Ra according to Ra/GPP ratio and GPP-NPP
  3. Heterotrophic respiration was estimated according to Rh = NPP - HerbComsum - Fire - Csink (NEP)
  4. Rs_topdown =  Rh + Rroot
  5. We then compare this top-down Rs with published Rs in the past decades (n=27)n to see how much probability they will match
  
## Questions and more anlysis in the future
* How unlikely is it that we can reconcile current Rs and GPP estimates? 
  E.g. what % of the distribution of each 'allows' us to match the central estimate of the other? (<30%)
* What is driving the differences between botton-up GPP and reported GPP and top-down Rs and reported Rs?
* How do we resolve this? Where is better/more data needed? 
* These questions all resolve around influence x uncertainty = importance, right? This is basically the 'Dietze graph'.
* Which is more likely, low GPP or high Rs? -- discussion
* Savanna and grassland more than 20% area, but sample numbers are too small for Frrot, RC, and Ra/Gpp ratio
* Ra < 0 how to resolve this (resolved)

## Importance/implication/benifits
* Provide another way to estimate GPP and Rs
* Provide a method to compare bottom-up GPP (start from field Rs/Froot/RC/Rroot measurements, then modelling to estimate GPP) with remote sensing or land model GPP.
* Provide a method to compare top-down Rs (Rs seperated from remote sensing land model based GPP) with filed experiment based Rs (field Rs measurement, then modeling).


```{r load packages, message=FALSE, echo=FALSE}
# Constants
OUTPUT_DIR		<- "outputs/"
SEPARATOR		<- "-------------------------------------------"
DATA_DIR <- 'data'
# Create output and log folders if they do not exist
if(!file.exists(OUTPUT_DIR)) dir.create(OUTPUT_DIR)

# install.packages('kableExtra')
library(kableExtra)
# Load required packages
# install.packages("data.table")
library(data.table)
# install.packages("lubridate")
library(lubridate)
# install.packages("cowplot")
library(cowplot)
library(knitr)
# install.packages("ggpubr")
library("ggpubr")
# install.packages("reshape")
library(reshape)
library(ggplot2)
# install.packages("ggmap")
library(ggmap)
# install.packages("maps")
library(maps)
# install.packages("mapdata")
library(mapdata)
library(tidyr)
# install.packages("hexbin")
library(hexbin)
library(dplyr)
# Source all needed functions
theme_set(theme_bw())
source('functions.R')
```



```{r self defined functions, echo=FALSE}
# prepare Ra-GPP ratio 
prep_RaGpp <- function() {
  # Ra/GPP
  RaGPP$Source <- "Meta-papers"
  RaGPP %>% select(Ecosystem, Leaf_habit, RaGPP_ratio, Source) -> sdata2
  srdb_v4 %>% 
    # select(Study_number, Biome, Ecosystem_type, Leaf_habit, Latitude, Longitude, Ra_annual, ER, GPP, NPP) %>% 
    select(Ecosystem_type, Leaf_habit, Rs_annual, Ra_annual, Rh_annual, ER, GPP, NPP) %>% 
    filter(!is.na(GPP) | !is.na(NPP) | !is.na(ER) ) -> sub_srdb
  sub_srdb$ER <- ifelse(is.na(sub_srdb$ER), sub_srdb$GPP, sub_srdb$ER) # assume NEP very small
  sub_srdb$RaGPP_ratio <- (sub_srdb$ER - sub_srdb$NPP)/sub_srdb$GPP
  sub_srdb %>% filter (RaGPP_ratio > 0 & RaGPP_ratio <1 & !is.na(Ecosystem_type)) -> sub_srdb
  sub_srdb$Source <- "srdb_v4"
  sub_srdb %>% select(Ecosystem_type, Leaf_habit, RaGPP_ratio, Source) -> sub_srdb
  colnames(sub_srdb) <- colnames(sdata2)
  
  sdata2 <- rbind(sdata2, sub_srdb)
  return (sdata2)
}

# plot froot
plot_froot <- function (sdata) {
  var_obs <- nrow(sdata)
  # plot Froot (root respiration/aototrophic respiration ratio) by ecosystem type
  Fl_plot <- ggplot(sdata, aes(x = IGBP2, y=Froot)) + geom_violin() +
    geom_jitter(shape=16, position=position_jitter(0.2), col = "gray") +
    geom_boxplot(width=.1) +
    ylab("Froot") +
    # stat_summary(fun.y=median, geom="point", size=2, color="red") +
    scale_x_discrete(limits = c("DF", "EF", "MF", "Other"),
                     labels = c("DF (n=13)","EF (n=91)", "MF (n=7)", "Other (n=11)") ) +
    theme(axis.title.x=element_blank()) +
    ylab("Rroot-Ra ratio")
  
  print(Fl_plot)
}
```



```{r preliminaries, echo=FALSE}
# message=FALSE, include=FALSE, echo=FALSE, cache=TRUE
# Set chunks defaults; these options will be applied to all subsequent chunks
knitr::opts_chunk$set(results = 'hide', message = TRUE, include = TRUE, echo = FALSE, warning = FALSE,
                      fig.height = 4, fig.width = 8, cache = TRUE)
```


## Method

```{r load data}
# input data
srdb_v4 <- read.csv(file.path(DATA_DIR, 'srdbv4.csv')) # SRDB_V4 data
FlFsFr <- read.csv(file.path(DATA_DIR, 'FlFsFrSummary.csv')) # Froot (Rroot/Rs) and Fshoot (Rshoot/Rs) from 39 papers
GPP <- read.csv(file.path(DATA_DIR, 'GlobalGPP_Sum.csv')) # 65 GPP estimates from published paper
GPP %>% filter(Year > 1980) -> GPP
NPP <- read.csv(file.path(DATA_DIR, 'ITONPP.CSV')) # 251 NPP estimates from ITO (2011)
GlobalRs <- read.csv(file.path(DATA_DIR, 'GlobalRs.csv')) # 27 estimates of global Rs
GlobalRs %>% filter(Year > 1980) -> GlobalRs
# (Ra/GPP) Piao, Shilong, et al. "Forest annual carbon cost: A global‐scale analysis of autotrophic respiration." Ecology 91.3 (2010): 652-661.
RaGPP <- read.csv(file.path(DATA_DIR, 'RaGPP.csv')) 
IGBP <- read.table(file.path(DATA_DIR, 'IGBP.txt'), header = TRUE, sep = ",")
# sort(unique(IGBP$IGBP2001Pr))
# There are 16 IGBP group
IGBP %>% mutate(IGBP = case_when( IGBP2001Pr == 0 ~ "WAT",
                                  IGBP2001Pr == 1 ~ "ENF",
                                  IGBP2001Pr == 2 ~ "EBF",
                                  IGBP2001Pr == 3 ~ "DNF",
                                  IGBP2001Pr == 4 ~ "DBF",
                                  IGBP2001Pr == 5 ~ "MF",
                                  IGBP2001Pr == 6 ~ "CSH",
                                  IGBP2001Pr == 7 ~ "OSH",
                                  IGBP2001Pr == 8 ~ "WSA",
                                  IGBP2001Pr == 9 ~ "SAV",
                                  IGBP2001Pr == 10 ~ "GRA",
                                  IGBP2001Pr == 11 ~ "WET",
                                  IGBP2001Pr == 12 ~ "CRO",
                                  IGBP2001Pr == 13 ~ "URB",
                                  IGBP2001Pr == 14 ~ "CVM",
                                  IGBP2001Pr == 15 ~ "SNO",
                                  IGBP2001Pr == 16 ~ "BSV",
                                  TRUE ~ "UNC" ) ) -> IGBP
# BBL: Interesting; I didn't know case_when! Aanother way to do this would be 
# IGBP_list <- c("WAT", "ENF", ...)
# IGBP$IGBP <- IGBP_list[IGBP$IGBP2001Pr]
# 16 IGBP group into 10 Ecosystems
IGBP %>% mutate (Ecosystem = case_when( IGBP2001Pr == 0 ~ "Water",
                                        IGBP2001Pr == 1 ~ "Forest",
                                        IGBP2001Pr == 2 ~ "Forest",
                                        IGBP2001Pr == 3 ~ "Forest",
                                        IGBP2001Pr == 4 ~ "Forest",
                                        IGBP2001Pr == 5 ~ "Forest",
                                        IGBP2001Pr == 6 ~ "Shrubland",
                                        IGBP2001Pr == 7 ~ "Shrubland",
                                        IGBP2001Pr == 8 ~ "Savanna",
                                        IGBP2001Pr == 9 ~ "Savanna",
                                        IGBP2001Pr == 10 ~ "Grassland",
                                        IGBP2001Pr == 11 ~ "Wetland",
                                        IGBP2001Pr == 12 ~ "Agriculture",
                                        IGBP2001Pr == 13 ~ "Urbland",
                                        IGBP2001Pr == 14 ~ "Agriculture",
                                        IGBP2001Pr == 15 ~ "Snow",
                                        IGBP2001Pr == 16 ~ "Desert",
                                        TRUE ~ "Unclassified")) -> IGBP

IGBP %>% mutate(IGBP_group = case_when( IGBP2001Pr == 0 ~ "WAT",
                                  IGBP2001Pr == 1 ~ "EF",
                                  IGBP2001Pr == 2 ~ "EF",
                                  IGBP2001Pr == 3 ~ "DF",
                                  IGBP2001Pr == 4 ~ "DF",
                                  IGBP2001Pr == 5 ~ "MF",
                                  IGBP2001Pr == 6 ~ "SH",
                                  IGBP2001Pr == 7 ~ "SH",
                                  IGBP2001Pr == 8 ~ "GRA",
                                  IGBP2001Pr == 9 ~ "GRA",
                                  IGBP2001Pr == 10 ~ "GRA",
                                  IGBP2001Pr == 11 ~ "WET",
                                  IGBP2001Pr == 12 ~ "CRO",
                                  IGBP2001Pr == 13 ~ "URB",
                                  IGBP2001Pr == 14 ~ "CRO",
                                  IGBP2001Pr == 15 ~ "SNO",
                                  IGBP2001Pr == 16 ~ "BSV",
                                  TRUE ~ "UNC" ) ) -> IGBP

# DGRsD <- read.csv(paste0(DATA_DIR,'/', 'DGRsD.csv'), header = TRUE)
AGB <- read.table(paste0(DATA_DIR,'/', 'GLC2000_Point.txt'), header = T, sep = ",")
AGB %>% filter(RASTERVALU != -9999) -> AGB

# head(AGB)
# max(AGB$RASTERVALU)
# min(AGB$RASTERVALU)
# 
# max(SOC$RASTERVALU)
# min(SOC$RASTERVALU)

SOC <- read.table(paste0(DATA_DIR,'/', 'SOC_Point.txt'), header = T, sep = ",")

srdb_v4 %>% 
  select(Study_number, Biome, Ecosystem_type, Leaf_habit, Latitude, Longitude, RC_annual, Ra_annual, Rh_annual) %>% 
  filter(RC_annual > 0 & RC_annual < 1) -> sub_srdb

# calculate Froot
Froot <- cal_Froot(FlFsFr, srdb_v4) 
# class(Froot$IGBP)
# check whether Froot + Fshoot = 1

Froot %>% mutate(IGBP2 = case_when( 
                                    IGBP %in% c('DBF', 'Deciduous') & Ecosystem == 'Forest' ~ "DF",
                                    IGBP %in% c('EBF', 'ENF', 'Evergreen') & Ecosystem == 'Forest' ~ "EF",
                                    IGBP %in% c('MF', 'Mixed') & Ecosystem == 'Forest' ~ "MF",
                                    TRUE ~ "Other"
                                    )) -> Froot
```


## Sites spatial distribution
```{r plot sites spatial distribution, echo=FALSE, warning=FALSE}
# plot sites distribution 
# Figure. Sites distribution -- FlFsFr sites and RC/HC sites
plot_sites(srdb_v4, FlFsFr)
```


## Ratio by vegetation types {.tabset .tabset-fade .tabset-pills}
### Plot GPP
```{r plot GPP and RSG, fig.height = 6, fig.width = 8}
# panel (a). all 65 GPP estimates
# panel (b). GPP vs year, the red dots in panel c are calculated trend from panel b (increase rate = 0.2 if outliers removed, increase rate = 0.42 if outlers kept)
# panel (c). 16 GPP estimates reported GPP increase trend
plot_GPP(GPP, GlobalRs)
ggsave("outputs/FigureS1.jpg", width = 8, height = 6, dpi = 300, units = "in")
```


### Plot Rroot-Rs ratio

```{r plot RC, fig.height=6, fig.width=8}
# Plot global Rs, NPP, Fire-burned Carbon, and RC by ecosystem (root respiration to soil respiration ratio)
# more details please see function plot_Rroot_Rs()
# prepare data first
sub_srdb %>% select(Leaf_habit) %>% group_by(Leaf_habit) %>% summarise(count = n())
sub_srdb %>% mutate(IGBP2 = case_when( Ecosystem_type == "Agriculture" ~ "Agriculture"
                                       , Ecosystem_type == "Grassland" ~ "Grassland"
                                       , Ecosystem_type %in% c("Savanna", "Shrubland") ~ "Shrubland"
                                       , Ecosystem_type == "Forest" & Leaf_habit == 'Deciduous' ~ "DF"
                                       , Ecosystem_type == "Forest" & Leaf_habit == 'Evergreen' ~ "EF"
                                       , Ecosystem_type == "Forest" & Leaf_habit == 'Mixed' ~ "Mixed"
                                       , TRUE ~ "Other" )) -> sub_srdb
sub_srdb %>% group_by(Ecosystem_type) %>% summarise(count = n())

plot_Rroot_Rs (GlobalRs, sub_srdb, NPP) 
ggsave("outputs/FigureS2.jpg", width = 8, height = 6, dpi = 300, units = "in")
```

### Plot Rroot-Ra ratio
```{r plot Rroot-Ra ratio, fig.height = 4, fig.width = 6}
# potential issue: only two points from grassland
# no point from tundra 
# BBL: difficult to see what's going on here; really needs more comments!
plot_froot (Froot)
ggsave("outputs/FigureS3.jpg", width = 6, height = 4, dpi = 300, units = "in")
```



### Plot Ra to GPP ratio
```{r plot Ra-GPP ratio}
# only one point from tundra
var_RaGpp <- prep_RaGpp()
var_RaGpp %>% count(Ecosystem)
var_RaGpp %>% count(Leaf_habit)

var_RaGpp %>% mutate(IGBP2 = case_when( Ecosystem %in% c( "Crop", "Tundra", "Wetland", "Agriculture" ) ~ "Other"
                                         , Ecosystem %in% c("Grassland", "Savanna") ~ "Grassland" 
                                         , Ecosystem == "Forest" & Leaf_habit == "Deciduous" ~ "Deciduous"
                                         , Ecosystem == "Forest" & Leaf_habit == "Evergreen" ~ "Evergreen"
                                         , Ecosystem == "Forest" & Leaf_habit == "Forest" ~ "Forest"
                                         , Ecosystem == "Forest" & Leaf_habit == "Mixed" ~ "Forest"
                                         , TRUE ~ "NOTA" )) -> var_RaGpp

var_RaGpp %>% count(IGBP2)
var_RaGpp %>% count(Ecosystem)

plot_RaGPP(var_RaGpp)

ggsave("outputs/FigureS4.jpg", width = 6, height = 4, dpi = 300, units = "in")
var_RaGpp %>% group_by (Ecosystem) %>% summarise ( count = n() ) %>% mutate (percent = count/sum(count)) %>%  arrange (desc(Ecosystem))
```


```{r calculate area percentage of each vegetation type}
# IGBP area by type
IGBP %>% count(IGBP_group)

IGBP %>% filter(IGBP_group != "WAT") %>% 
  group_by(IGBP_group) %>% 
  summarise(count = n()) %>% 
  mutate(percent = count / sum(count)) -> IGBP_summary

ag_area <- IGBP_summary[IGBP_summary$IGBP_group == "CRO",]$percent
df_area <- IGBP_summary[IGBP_summary$IGBP_group == "DF",]$percent
ef_area <- IGBP_summary[IGBP_summary$IGBP_group == "EF",]$percent
mf_area <- IGBP_summary[IGBP_summary$IGBP_group == "MF",]$percent
gra_area <- IGBP_summary[IGBP_summary$IGBP_group == "GRA",]$percent
shr_area <- IGBP_summary[IGBP_summary$IGBP_group == "SH",]$percent

```


```{r prepare variables for boostrapping}
# 1. prepare RC (Rroot-to-Rs ratio data) **********
# get subset data for Bottom-up approache to estimate GPP, seperate into 7 groups
var_RC <- sub_srdb %>% select(RC_annual, Ecosystem_type, IGBP2) %>% filter(RC_annual > 0.01 & RC_annual < 0.99 & !is.na(RC_annual)) # dataset contain RC records
var_for_rc <- var_RC %>% filter (Ecosystem_type == "Forest") # dataset contain RC records only for Forest ecosystem
# forest further seperate into 3 groups
var_df_rc <- var_RC %>% filter (IGBP2 == "DF") # dataset contain RC records only for DF
var_ef_rc <- var_RC %>% filter (IGBP2 == "EF") # dataset contain RC records only for EF ecosystem
var_mf_rc <- var_RC %>% filter (IGBP2 == "Mixed") # dataset contain RC records only for MF ecosystem
var_agr_rc <- var_RC %>% filter (Ecosystem_type == "Agriculture") # dataset contain RC records only for agriculture ecosystem
var_gra_rc <- var_RC %>% filter (Ecosystem_type == "Grassland") # dataset contain RC records only for grassland ecosystem
var_shr_rc <- var_RC %>% filter (Ecosystem_type %in% c("Shrubland", "Savanna") ) # dataset contain RC records only for grassland ecosystem
# Other groups
var_other_rc <- var_RC %>% filter (Ecosystem_type %in% c("Desert", "Wetland" )) # dataset contain RC records only for Other ecosystem
var_rest_rc <- var_RC # for those ecosystem (e.g., tundra) has no data, use all from RC dataset

# 2. prepare Froot (Rroot-to-Ra ratio data) **********
var_agr_froot = Froot[Froot$Ecosystem == "Agriculture",]$Froot # samples of Froot for agriculture
var_for_froot = Froot[Froot$Ecosystem == "Forest",]$Froot # samples of Froot for forest
var_gra_froot = Froot[Froot$Ecosystem == "Grassland",]$Froot # samples of Froot for grassland
var_sav_froot = Froot[Froot$Ecosystem == "Savanna",]$Froot # samples of Froot for savanna
var_wet_froot = Froot[Froot$Ecosystem == "Wetland",]$Froot # samples of Froot for agriculture
var_rest_froot = Froot$Froot # for those ecosystems has no data, use all data from Froot
# only seperate into ef and the rest
var_ef_froot = Froot[Froot$IGBP2 == "EF",]$Froot # Froot data for ef
var_df_froot = Froot[Froot$IGBP2 == "DF",]$Froot # Froot data for df
var_mf_froot = Froot[Froot$IGBP2 == "MF",]$Froot # Froot data for mf
var_other_froot = Froot[Froot$IGBP2 == "Other",]$Froot # samples of Froot for non-ef

# var_rest <- var_RC %>% filter (Ecosystem_type != "Forest") # dataset contain RC records for non-forest ecosystem
var_Rs <- GlobalRs %>% filter(!is.na(Rs) & is.na(SD) ) # dataset contain global Rs estimates (without standard diviation information)
var_GPP <- GPP[is.na(GPP$SD),]$GPP

```


## Bootstrapping GPP
```{r using boostrapping prepare data for GPP bottom-up and top-down comparison}
# Bottom-up approache to stimate GPP
# Rroot = Rs * RC (Rroot to Rs ratio, from srdb)
# Froot = Rroot / Ra (root autotrophic respiration to total autotrophic respiration fraction/ratio, calculation please see plot_froot function)
# Fshoot = 1-Froot
# FsFr = (1-Froot)/Froot (FsFr = Rshoot/Rroot ratio)
# Ra = Rroot + Rshoot
# GPP = NPP + Rroot + Rshoot

n_samp <- 50000 # take 50000 samples
# n_rep <- 50000 

# bootstrapping
set.seed(20190827)

gpp_results <- tibble(
  #Step 1. Prepare Global Rs data
  # Rs = replicate( n_rep, sample(var_Rs$Rs, n_samp, replace = TRUE) %>% mean() ),
  # Rs1 = rnorm(n = n_samp, mean = 68, sd = 4), # not sure about the SD = 4
  # for those without SD Rs, randomly take a sample from global Rs estimates, with replace=TRUE, repeat 50000 times
  Rs1 = sample(var_Rs$Rs, n_samp, replace = TRUE), 
  # for those studies reported Rs (mean and SD, 9 studies) using rnorm functions
  Rs2 = rnorm(n = n_samp, mean = 80.4, sd = 16.9), # for those with standard diviation, 50000 samples were take from normal distribution ~ (mean=80.4, sd=16.9)
  # repeat until all Rs estimates with sd samples were taken and hold in the outputs
  Rs3 = rnorm(n = n_samp, mean = 98, sd = 12), # same as above
  Rs4 = rnorm(n = n_samp, mean = 79, sd = 7.65), # same as above
  Rs5 = rnorm(n = n_samp, mean = 94.4, sd = 4.59), # same as above
  Rs6 = rnorm(n = n_samp, mean = 91, sd = 2.05), # same as above
  Rs7 = rnorm(n = n_samp, mean = 94.3, sd = 9.13), # same as above
  Rs8 = rnorm(n = n_samp, mean = 78.34, sd = 2.25), # same as above
  Rs9 = rnorm(n = n_samp, mean = 72.55, sd = 7.13), # same as above
  Rs10 = rnorm(n = n_samp, mean = 93.3, sd = 3.11), # same as above
  # Rs was calculated as the mean of above 10 subsamples
  Rs = (Rs1 + Rs2 + Rs3 + Rs4 + Rs5 + Rs6 + Rs7 + Rs8 + Rs9 + Rs10)/10,
  
  # Step 2. Prepare NPP, boosting from ITO paper
  NPP = sample(NPP$NPP, n_samp, replace = TRUE) , # take 500000 samples from NPP records
  
  # Step 3. Prepare GPP (same logic as global Rs)
  gpp1 = rnorm(n = n_samp, mean = 124.6, sd = 2.7), # if gpp have mean and standard diviation, take 50000 samples from the normal distribution
  gpp2 = rnorm(n = n_samp, mean = 135.7, sd = 21.712), # same as above
  gpp3 = rnorm(n = n_samp, mean = 109.29, sd = 27.33), # same as above
  gpp4 = rnorm(n = n_samp, mean = 118, sd = 26), # same as above
  gpp5 = rnorm(n = n_samp, mean = 119, sd = 6), # same as above
  gpp6 = rnorm(n = n_samp, mean = 110.5, sd = 21.3), # same as above
  gpp7 = rnorm(n = n_samp, mean = 117, sd = 13), # same as above
  gpp8 = rnorm(n = n_samp, mean = 147, sd = 16), # same as above
  gpp9 = rnorm(n = n_samp, mean = 119, sd = 6), # same as above
  gpp10 = rnorm(n = n_samp, mean = 122, sd = 25), # same as above
  # for those do not have SD information, take 50000 samples using boosting 
  gpp11 = sample(var_GPP, n_samp, replace = TRUE), 
  # calculate GPP, this is the Top-dowm GPP (collected from previous studies, and resampling)
  GPP_raw =  (gpp1 + gpp2 + gpp3 + gpp4 + gpp5 + gpp6 + gpp7 + gpp8 + gpp9 + gpp10 + gpp11)/11, 
  
  # Step 4. sampling Rroot-to Rs ratio (RC)
  Rc_agr = sample(var_agr_rc$RC_annual, n_samp, replace = TRUE), # randomly take a sample from RC estimates for agriculture, repeat 50000 times
  # forest further seperate into 3 groups
  Rc_df = sample(var_df_rc$RC_annual, n_samp, replace = TRUE), # randomly take a sample from RC estimates for df, repeat 50000 times
  Rc_ef = sample(var_ef_rc$RC_annual, n_samp, replace = TRUE), # randomly take a sample from RC estimates for df, repeat 50000 times
  Rc_mf = sample(var_mf_rc$RC_annual, n_samp, replace = TRUE), # randomly take a sample from RC estimates for df, repeat 50000 times
  # other ecosystems
  Rc_gra = sample(var_gra_rc$RC_annual, n_samp, replace = TRUE), # randomly take a sample from RC estimates for grassland, repeat 50000 times
  Rc_shr = sample(var_shr_rc$RC_annual, n_samp, replace = TRUE), # randomly take a sample from RC estimates for grassland, repeat 50000 times
  # note that for the rest ecosystems we used samples from all RC records rather that from other groups
  Rc_rest = sample(var_rest_rc$RC_annual, n_samp, replace = TRUE), 
  
  
  # Step 5. Sampling Rroot-to-Ra ratio (Froot)
  Froot_df = sample(var_df_froot, n_samp, replace = TRUE),
  Froot_ef = sample(var_ef_froot, n_samp, replace = TRUE),
  Froot_mf = sample(var_mf_froot, n_samp, replace = TRUE),
  # note that for the rest ecosystems we used samples from all RC records rather that from other groups
  Froot_rest = sample(var_rest_froot, n_samp, replace = TRUE),
  
  # Step 6. calculation ********************
  # Rc = replicate( n_rep, var_RC$RC_annual, n_samp, replace = TRUE) %>% mean() ), # senario 0: Rc not seperate into different ecosystems, not used
  Rc = sample(var_RC$RC_annual, n_samp, replace = TRUE),
  # Rc = Rc_df*df_area + Rc_ef*ef_area + Rc_mf*mf_area + Rc_rest*1-(df_area + ef_area + mf_area), 
  
  # Senario 2: Rc seperate into 4 groups: agriculture, forest, grassland, and rest (all other ecosystems with obs<40, take samples from all RC records)
  # only seperate into four groups, agriculture, forest, grassland and the rest
  # 1-(0.191 + 0.214 + 0.153 + 0.268)
  Rc2 = (Rc_agr*ag_area + Rc_df*df_area + Rc_ef*ef_area + Rc_mf*mf_area + Rc_gra*gra_area + Rc_shr*shr_area 
         + Rc_rest*(1-ag_area-df_area-ef_area-mf_area-gra_area-shr_area)), 
  # mean(sample(var_for$RC_annual, n_samp, replace = TRUE))
  # sample(var_df$RC_annual, n_samp, replace = TRUE) %>% mean
  # sample(var_ef$RC_annual, n_samp, replace = TRUE) %>% mean
  # sample(var_mf$RC_annual, n_samp, replace = TRUE) %>% mean
    
  # senario1: Froot not seperate into different ecosystems
  Froot = sample(Froot$Froot, n_samp, replace = TRUE) ,
  
  FsFr = (1 - Froot) / Froot, # calculate Froot to Fshoot ratio of senario 1
  # Bottom up estimate of GPP, senario 1
  Rroot = Rs * Rc, # calculate root respiration
  Rshoot = Rroot * FsFr, # calculate shoot respiration
  
  # Senario 2: seperate into 2 ecosystems and weighted by area
  # Froot %>% count(Ecosystem)
  Froot2 = Froot_df*df_area + Froot_ef*ef_area + Froot_mf*mf_area + Froot_rest * (1-df_area-ef_area-mf_area),
  FsFr2 = (1 - Froot2) / Froot2, # calculate Froot to Fshoot ratio of senario 2
  # # Bottom up estimate of GPP, senario 2
  Rroot2 = Rs * Rc2,
  Rshoot2 = Rroot2 * FsFr2,
  
  GPP = NPP + Rroot + Rshoot, # scenario 1: calculate GPP
  GPP2 = NPP + Rroot2 + Rshoot2 # scenario 2: calculate GPP2
)
```


## GPP results diagnose {.tabset .tabset-fade .tabset-pills}
### senario 1 - ratio __not__ weighted by ecosystem
```{r plot GPP comparison}
# hist(gpp_results$GPP) + abline (v = quantile (gpp_results$GPP, c(0.025, 0.5, 0.975)), col = "red", lty = "dashed")
# mean(gpp_results$GPP)
# max(gpp_results$GPP)

# senario 1
gpp_qt <- quantile(gpp_results$GPP, c(0.025, 0.5, 0.975)) %>% round(0)
gpp_qt_raw <- quantile(gpp_results$GPP_raw, c(0.025, 0.5, 0.975)) %>% round(0)
gpp_results %>% ggplot() + aes(GPP) + 
  # geom_histogram(col = "black", fill = "gray", bins = 30) + 
  geom_density(stat = 'density', col = "gray", fill = rgb(1,0,1, alpha = 0.5) ) +
  # geom_histogram(aes(GPP_raw), col = "blue", fill = "blue", bins = 30, alpha = 0.35) +
  geom_density(aes(GPP_raw), stat = "density", col = "gray", fill = rgb(0,1,1, alpha = 0.5)) +
  # geom_vline(xintercept = c(quantile (gpp_results$GPP_raw, c(0.025, 0.975))),col = "gray", linetype = 2, size = 1) +
  annotate("text", x = 150, y = 0.035, label = paste0("2.5%, 50%, 97.5% quantile (", gpp_qt[1],", ", gpp_qt[2], ", ", gpp_qt[3], ")")
           , hjust = 0, col = rgb(1,0,1)) +
  annotate("text", x = 150, y = 0.025, label = paste0("2.5%,50%,97.5% quantile (", gpp_qt_raw[1],", ", gpp_qt_raw[2], ", ", gpp_qt_raw[3], ")")
           , hjust = 0, col = rgb(0,1,1)) +
  xlab(expression(GPP~"("~Pg~C~yr^{-1}~")"))

# ggsave("outputs/Figure2.jpg", width = 6, height = 4, dpi = 300, units = "in")

```

### Sematio 2 - ratio __weighted__ by ecosystem
```{r}
# senario 2
rsg_qt <- quantile(gpp_results$GPP2, c(0.025, 0.5, 0.975)) %>% round(0)
gpp_qt_raw <- quantile(gpp_results$GPP_raw, c(0.025, 0.5, 0.975)) %>% round(0)
ggplot(gpp_results) + aes(GPP2) + 
  # geom_histogram(col = "black", fill = "gray", bins = 30) + 
  geom_density(stat = 'density', col = "gray", fill = rgb(1,0,1, alpha = 0.5) ) +
  # geom_histogram(aes(GPP_raw), col = "blue", fill = "blue", bins = 30, alpha = 0.35) +
  geom_density(aes(GPP_raw), stat = "density", col = "gray", fill = rgb(0,1,1, alpha = 0.5)) +
  # geom_vline(xintercept = c(quantile (gpp_results$GPP_raw, c(0.025, 0.975))),col = "gray", linetype = 2, size = 1) +
  annotate("text", x = 150, y = 0.035, label = paste0("2.5%, 50%, 97.5% quantile (", rsg_qt[1],", ", rsg_qt[2], ", ", rsg_qt[3], ")")
           , hjust = 0, col = rgb(1,0,1)) +
  annotate("text", x = 150, y = 0.025, label = paste0("2.5%,50%,97.5% quantile (", gpp_qt_raw[1],", ", gpp_qt_raw[2], ", ", gpp_qt_raw[3], ")")
           , hjust = 0, col = rgb(0,1,1)) +
  xlab(expression(GPP~"("~Pg~C~yr^{-1}~")"))
```

### GPP vs components relationships
```{r}
gpp_results %>% ggplot() + aes(Rc) + 
  # geom_histogram(col = "black", fill = "gray", bins = 30) + 
  geom_density(stat = 'density', col = "gray", fill = rgb(1,0,1, alpha = 0.5) )

gpp_results %>% ggplot() + aes(Rc2) + 
  # geom_histogram(col = "black", fill = "gray", bins = 30) + 
  geom_density(stat = 'density', col = "gray", fill = rgb(1,0,1, alpha = 0.5) )

gpp_results %>% ggplot() + aes(Rroot) + 
  # geom_histogram(col = "black", fill = "gray", bins = 30) + 
  geom_density(stat = 'density', col = "gray", fill = rgb(1,0,1, alpha = 0.5) )

gpp_results %>% ggplot() + aes(Rroot2) + 
  # geom_histogram(col = "black", fill = "gray", bins = 30) + 
  geom_density(stat = 'density', col = "gray", fill = rgb(1,0,1, alpha = 0.5) )

# GPP ~ NPP
ggplot(gpp_results) +
  aes(x = NPP, y = GPP2) +
  geom_hex() +
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") + 
  xlab(expression(NPP~"("~Pg~C~yr^{-1}~")")) +
  ylab(expression(GPP~"("~Pg~C~yr^{-1}~")"))

# GPP ~ Rs
ggplot(gpp_results) +
  aes(x = gpp_results$Rs, y = GPP2) +
  geom_hex() +
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") +
  xlab(expression(R[S]~"("~Pg~C~yr^{-1}~")")) +
  ylab(expression(GPP~"("~Pg~C~yr^{-1}~")"))

# GPP ~ Rc
ggplot(gpp_results) +
  aes(x = Rc, y = GPP2) +
  geom_hex() +
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") +
  xlab(expression(RC)) +
  ylab(expression(GPP~"("~Pg~C~yr^{-1}~")"))

# GPP ~ Froot
ggplot(gpp_results) +
  aes(x = Froot, y = GPP2) +
  geom_hex() +
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") +
  xlab(expression(Froot)) +
  ylab(expression(GPP~"("~Pg~C~yr^{-1}~")"))

# GPP ~ Fshoot
ggplot(gpp_results) +
  aes(x = 1-Froot, y = GPP2) +
  geom_hex() +
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") +
  xlab(expression(Fshoot)) +
  ylab(expression(GPP~"("~Pg~C~yr^{-1}~")"))
```

### Probability of aggrement
```{r compute propotion of aggrement between top-down RSG and bottom-up GPP}
# Quantify the probability GPP following (107, 133)
# about 29% chance GPP will between (107-133)
y <- mean(gpp_results$GPP)
s <- sd(gpp_results$GPP)
# accume GPP following normal distribution, with mean = y and sd = s, we can calculate the properbility of GPP following (107-133) 95%CI
p_gpp_high <- pnorm ((gpp_qt_raw[3] - y) / (s))
p_gpp_low <- pnorm ((gpp_qt_raw[1] - y) / (s))
p_gpp1 <- (p_gpp_high - p_gpp_low) %>% round(4)
print(paste0("The probability of aggrement in scenario 1 is: ", p_gpp1))
# pnorm (-0.329)

# senario 2 looks like a little bit right tail skew
# about 18%
y <- mean(gpp_results$GPP2)
s <- sd(gpp_results$GPP2)
p_gpp_high <- pnorm ((gpp_qt_raw[3] - y) / (s))
p_gpp_low <- pnorm ((gpp_qt_raw[1] - y) / (s))
p_gpp2 <- (p_gpp_high - p_gpp_low) %>% round(4)
print(paste0("The probability of aggrement in scenario 2 is: ", p_gpp2))
```

The probability of aggrement in scenario 1 is ```r p_gpp1*100``` %.  
The probability of aggrement in scenario 2 is ```r p_gpp2*100``` %.


```{r prepare variables fro RSG boostrapping}
# Froot used the same data as bottom-up approach
var_Fire <- c(2, 3.5, 7.3, 4, 5.1, 2.02, 2.71, 3.02, 2.08)
# prepare RaGPP ratio data
var_RaGpp %>% count(IGBP2)
var_RaGpp_df = var_RaGpp[var_RaGpp$IGBP2 == "Deciduous", ]$RaGPP_ratio
var_RaGpp_ef = var_RaGpp[var_RaGpp$IGBP2 == "Evergreen", ]$RaGPP_ratio
# mix forest are from all forest due to lack of data
var_RaGpp_mf = var_RaGpp[var_RaGpp$Ecosystem == "Forest", ]$RaGPP_ratio 
var_RaGpp_gra = var_RaGpp[var_RaGpp$Ecosystem == "Grassland", ]$RaGPP_ratio

```

## Bootstrapping RSG
```{r using boostrap prepare data for RSG bottom-up and top-down comparison}
# topdowm approach to eatimate Rs
# Ra = GPP * Ra_GPP_ratio (calculated based on data from a summary paper and from srdb, see plot_RaGPP function)
# Ra = GPP - NPP (two approaches yield a very similar results, we thus used the average of two estimates)
n_samp <- 50000

set.seed(20190827)
rs_results <- tibble(
  # Step 1. Prepare GPP data
  gpp1 = rnorm(n = n_samp, mean = 124.6, sd = 2.7), # if gpp have mean and standard diviation, take 50000 samples from the normal distribution
  gpp2 = rnorm(n = n_samp, mean = 135.7, sd = 21.712), # same as above
  gpp3 = rnorm(n = n_samp, mean = 109.29, sd = 27.33), # same as above
  gpp4 = rnorm(n = n_samp, mean = 118, sd = 26), # same as above
  gpp5 = rnorm(n = n_samp, mean = 119, sd = 6), # same as above
  gpp6 = rnorm(n = n_samp, mean = 110.5, sd = 21.3), # same as above
  gpp7 = rnorm(n = n_samp, mean = 117, sd = 13), # same as above
  gpp8 = rnorm(n = n_samp, mean = 147, sd = 16), # same as above
  gpp9 = rnorm(n = n_samp, mean = 119, sd = 6), # same as above
  gpp10 = rnorm(n = n_samp, mean = 122, sd = 25), # same as above
  # for those do not have SD information, take 50000 samples using boosting 
  gpp11 = sample(var_GPP, n_samp, replace = TRUE), 
  # calculate GPP, this is the Top-dowm GPP (collected from previous studies, and resampling)
  GPP_raw =  (gpp1 + gpp2 + gpp3 + gpp4 + gpp5 + gpp6 + gpp7 + gpp8 + gpp9 + gpp10 + gpp11)/11, 
  
  # Step 2. prepare other components, NPP and NPP consumed by fire etc.
  NPP = sample(NPP$NPP, n_samp, replace = TRUE),
  HerbComsum = rnorm(n = n_samp, mean = 2.2, sd = 0.2),
  Fire = sample(var_Fire, n_samp, replace = TRUE),
  sink = rnorm(n = n_samp, mean = 2.10, sd = 0.28),
  
  # Step 3. prepare Rs data
  Rs1 = sample(var_Rs$Rs, n_samp, replace = TRUE),
  Rs2 = rnorm(n = n_samp, mean = 80.4, sd = 16.9),
  Rs3 = rnorm(n = n_samp, mean = 98, sd = 12),
  Rs4 = rnorm(n = n_samp, mean = 79, sd = 7.65),
  Rs5 = rnorm(n = n_samp, mean = 94.4, sd = 4.59),
  Rs6 = rnorm(n = n_samp, mean = 91, sd = 2.05),
  Rs7 = rnorm(n = n_samp, mean = 94.3, sd = 9.13),
  Rs8 = rnorm(n = n_samp, mean = 78.34, sd = 2.25),
  Rs9 = rnorm(n = n_samp, mean = 72.55, sd = 7.13),
  Rs10 = rnorm(n = n_samp, mean = 93.3, sd = 3.11),
  Rs_raw = (Rs1 + Rs2 + Rs3 + Rs4 + Rs5 + Rs6 + Rs7 + Rs8 + Rs9 + Rs10)/10,
  
  # Step 4. Ra-to-GPP ratio boosting
  # RaGpp = sample(var_RaGpp$RaGPP_ratio, n_samp, replace = TRUE),
  RaGpp_df = sample(var_RaGpp_df, n_samp, replace = TRUE),
  RaGpp_ef = sample(var_RaGpp_ef, n_samp, replace = TRUE),
  RaGpp_mf = sample(var_RaGpp_mf, n_samp, replace = TRUE),
  RaGpp_gra = sample(var_RaGpp_gra, n_samp, replace = TRUE),
  # note that for the rest ecosystems, we used samples from all Froot rather than Other group
  RaGpp_rest = sample(var_RaGpp$RaGPP_ratio, n_samp, replace = TRUE),
  
  # Step 5. Sampling Rroot-to-Ra ratio (Froot)
  Froot_df = sample(var_df_froot, n_samp, replace = TRUE),
  Froot_ef = sample(var_ef_froot, n_samp, replace = TRUE),
  Froot_mf = sample(var_mf_froot, n_samp, replace = TRUE),
  Froot_gra = sample(var_gra_froot, n_samp, replace = TRUE),
  # note that for the rest ecosystems, we used samples from all Froot rather than Other group
  Froot_rest = sample(var_rest_froot, n_samp, replace = TRUE), 
  
  # Calculation **********
  # senario 1
  RaGpp = RaGpp_rest,
  # # senario 2
  RaGpp2 = RaGpp_df*df_area + RaGpp_ef*ef_area + RaGpp_mf*mf_area + RaGpp_gra*gra_area + RaGpp_rest*(1-df_area-ef_area-mf_area-gra_area),
  # 
  # # # get Froot (Froot2 is better, from bottom-up section, see above results)
  Froot2 = Froot_df*df_area + Froot_ef*ef_area + Froot_mf*mf_area + Froot_rest * (1-df_area-ef_area-mf_area),
  # 
  Ra1 = GPP_raw * RaGpp, # first way to calculate Ra, Ra = GPP * RaGPP
  Ra2 = GPP_raw - NPP, # second way to calculate Ra, Ra = GPP - NPP
  # # BBL: what's this comma? Why doesn't it cause an error?
  # # response: because few Ra2 < 0 (when GPP < NPP), for those estimates, just drop them? or use Ra1 replace them? or keep them?
  Ra3 = ifelse(Ra2 < 0, Ra1, Ra2),
  Ra_avg1 = (Ra1 + Ra3) / 2,
  # # 
  Rroot = Ra_avg1 * Froot2,
  Rshoot = Ra_avg1 * (1 - Froot2),
  Rs_topdown = NPP - HerbComsum - Fire - sink + Rroot,
  # 
  # # senario2
  Ra4 = GPP_raw * RaGpp2, # first way to calculate Ra, Ra = GPP * RaGPP
  Ra_avg2 = (Ra4 + Ra3) / 2,
  Rroot2 = Ra_avg2 * Froot2,
  Rshoot2 = Ra_avg2 * (1 - Froot2),
  Rs_topdown2 = NPP - HerbComsum - Fire - sink + Rroot2
)
```


## RSG results diagnose {.tabset .tabset-fade .tabset-pills}
### Scenario 1 - ratio __not__ weighted by ecosystem
```{r plot RSG results}
# mean(rs_results$Rs)
# max(rs_results$Rs)

# scenatio 1. RaGPP ratio not seperated
rsg_qt <- quantile(rs_results$Rs_topdown, c(0.025, 0.5, 0.975)) %>% round(0)
rsg_qt_raw <- quantile(rs_results$Rs_raw, c(0.025, 0.5, 0.975)) %>% round(0)
ggplot(rs_results) + aes(Rs_topdown) + 
  geom_density(col = "gray", fill = rgb(1,0,1, alpha = 0.25)) + 
  geom_density(aes(Rs_raw), col = "gray", fill = rgb(0,1,1, alpha = 0.5)) +
  # geom_vline(xintercept = c(quantile (rs_results$Rs_raw, c(0.025, 0.5, 0.975))),col = "red", linetype = "dotted", size = 1.5) +
  annotate("text", x = 100, y = 0.04, label = paste0("2.5%, 50%, 97.5% quantile (", rsg_qt[1],", ", rsg_qt[2], ", ", rsg_qt[3], ")")
           , hjust = 0, col = rgb(1,0,1)) +
  annotate("text", x = 100, y = 0.03, label = paste0("2.5%, 50%, 97.5% quantile (", rsg_qt_raw[1],", ", rsg_qt_raw[2], ", ", rsg_qt_raw[3], ")")
           , hjust = 0, col = rgb(0,1,1)) +
  ylab(expression(count))

# Rs raw data plot
# rsg_qt <- quantile(rs_results$Rs_raw, c(0.025, 0.5, 0.975)) %>% round(0)
# plot_rs_raw <- ggplot(rs_results) + aes(Rs_raw) + geom_histogram(col = "black", fill = "gray", bins = 30) + 
#   geom_vline(xintercept = c(quantile (rs_results$Rs, c(0.025, 0.5, 0.975))),col = "red", linetype = "dotted", size = 1.5) +
#   annotate("text", x = quantile(rs_results$Rs, 0.98), y = 5000, label = paste0("95%CI (", rsg_qt[1],", ", rsg_qt[2], ", ", rsg_qt[3], ")"), hjust = 0)

```

### Scenario 2 - ratio __weighted__ by ecosystem
```{r}
# senario 2. RaGPP ratio seperated into 2 groups
rsg_qt <- quantile(rs_results$Rs_topdown2, c(0.025, 0.5, 0.975)) %>% round(0)
rsg_qt_raw <- quantile(rs_results$Rs_raw, c(0.025, 0.5, 0.975)) %>% round(0)
ggplot(rs_results) + aes(Rs_topdown2) + 
  geom_density(col = "gray", fill = rgb(1,0,1, alpha = 0.25)) + 
  geom_density(aes(Rs_raw), col = "gray", fill = rgb(0,1,1, alpha = 0.5)) +
  # geom_vline(xintercept = c(quantile (rs_results$Rs_raw, c(0.025, 0.5, 0.975))),col = "red", linetype = "dotted", size = 1.5) +
  annotate("text", x = 100, y = 0.04, label = paste0("2.5%, 50%, 97.5% quantile (", rsg_qt[1],", ", rsg_qt[2], ", ", rsg_qt[3], ")")
           , hjust = 0, col = rgb(1,0,1)) +
  annotate("text", x = 100, y = 0.03, label = paste0("2.5%, 50%, 97.5% quantile (", rsg_qt_raw[1],", ", rsg_qt_raw[2], ", ", rsg_qt_raw[3], ")")
           , hjust = 0, col = rgb(0,1,1)) +
  ylab(expression(count))
```

### RSG and components relationship
```{r}
# Ra1 ~ Ra2
ggplot(rs_results) +
  aes(x = Ra1) + geom_histogram(col = "black", fill = "gray", bins = 30) +
  geom_histogram(aes(Ra3), col = "blue", fill = "skyblue", bins = 30, alpha = 0.5) +
  xlab(expression(Ra1~or~Ra3~"("~Pg~C~yr^{-1}~")")) +
  ylab(expression(count))

# Rs ~ GPP
ggplot(rs_results) +
  aes(x = GPP_raw, y = Rs_topdown) +
  geom_hex() +
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") +
  xlab(expression(GPP~"("~Pg~C~yr^{-1}~")")) +
  ylab(expression(R[S]~"("~Pg~C~yr^{-1}~")"))

# Rs ~ NPP
ggplot(rs_results) +
  aes(x = NPP, y = Rs_topdown) +
  geom_hex() +
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") +
  xlab(expression(NPP~"("~Pg~C~yr^{-1}~")")) +
  ylab(expression(R[S]~"("~Pg~C~yr^{-1}~")"))

# Rs ~ RaGpp
ggplot(rs_results) +
  aes(x = RaGpp, y = Rs_topdown) +
  geom_hex() +
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") +
  xlab(expression("Ra/GPP"~"("~Pg~C~yr^{-1}~")")) +
  ylab(expression(R[S]~"("~Pg~C~yr^{-1}~")"))

# Rs ~ Froot
ggplot(rs_results) +
  aes(x = Froot2, y = Rs_topdown) +
  geom_hex() +
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") +
  xlab(expression(Froot)) +
  ylab(expression(R[S]~"("~Pg~C~yr^{-1}~")"))

# Rs ~ Fire
ggplot(rs_results) +
  aes(x = Fire, y = Rs_topdown) +
  geom_hex() +
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") +
  xlab(expression(Fire~burned~"("~Pg~C~yr^{-1}~")")) +
  ylab(expression(R[S]~"("~Pg~C~yr^{-1}~")"))
```

### RSG probability of aggrement

```{r compute propotion of aggrement between top-down RSG and bottom-up RSG}
# accume Rs following normal distribution, with mean = y and sd = s, we can calculate the properbility of Rs following (80-92) 95%CI
# about 28% chance Rs will be between (80-92)
y <- mean(rs_results$Rs_topdown)
s <- sd(rs_results$Rs_topdown)
p_Rs_high <- pnorm ((y - rsg_qt_raw[1])/(s))
p_Rs_low <- pnorm ((y - rsg_qt_raw[3])/(s))
p_Rs1 <- (p_Rs_high - p_Rs_low) %>% round(4)
print(paste0("The probability of aggrement in scenario 1 is: ", p_Rs1))

# Senario 2: about 23% chance Rs will be between (80-92)
y <- mean(rs_results$Rs_topdown2)
s <- sd(rs_results$Rs_topdown2)
p_Rs_high <- pnorm ((y - rsg_qt_raw[1])/(s))
p_Rs_low <- pnorm ((y - rsg_qt_raw[3])/(s))
p_Rs2 <- (p_Rs_high - p_Rs_low) %>% round(4)
print(paste0("The probability of aggrement in scenario 2 is: ", p_Rs2))
```

The probability of aggrement in scenario 1 is ```r p_Rs1*100``` %.  
The probability of aggrement in scenario 2 is ```r p_Rs2*100``` %.


## Topdown and bottomup comparison - Figure 1
```{r combine GPP and RSG comparison into figure 2, fig.height=8, fig.width=6}
# senario 2, RC seperate into 4 groups, and Froot seperate into 2 groups
gpp_qt <- quantile(gpp_results$GPP2, c(0.025, 0.5, 0.975)) %>% round(0)
gpp_qt_raw <- quantile(gpp_results$GPP_raw, c(0.025, 0.5, 0.975)) %>% round(0)
ggplot(gpp_results) + aes(GPP2) + 
  geom_density(stat = 'density', col = "gray", fill = rgb(1,0,1, alpha = 0.5) ) +
  geom_density(aes(GPP_raw), stat = "density", col = "gray", fill = rgb(0,1,1, alpha = 0.5)) +
  # geom_vline(xintercept = c(quantile (gpp_results$GPP_raw, c(0.025, 0.5, 0.975))),col = "red", linetype = "dotted", size = 1.5) +
  annotate("text", x = 150, y = 0.035, label = paste0("2.5%, 50%, 97.5% quantile (", gpp_qt[1],", ", gpp_qt[2], ", ", gpp_qt[3], ")")
           , hjust = 0, col = rgb(1,0,1)) +
  annotate("text", x = 150, y = 0.025, label = paste0("2.5%,50%,97.5% quantile (", gpp_qt_raw[1],", ", gpp_qt_raw[2], ", ", gpp_qt_raw[3], ")")
           , hjust = 0, col = rgb(0,1,1)) +
  # annotate("text", x = 150, y = 0.045, label = expression("Rroot-to-R"[S]~"seperated into 4 groups")
  #          , hjust = 0) +
  xlab(expression(GPP~"("~Pg~C~yr^{-1}~")")) -> p1

# senario two
rsg_qt <- quantile(rs_results$Rs_topdown2, c(0.025, 0.5, 0.975)) %>% round(0)
rsg_qt_raw <- quantile(rs_results$Rs_raw, c(0.025, 0.5, 0.975)) %>% round(0)
ggplot(rs_results) + aes(Rs_topdown2) + 
  geom_density(col = "gray", fill = rgb(1,0,1, alpha = 0.5)) + 
  geom_density(aes(Rs_raw), col = "gray", fill = rgb(0,1,1, alpha = 0.5)) +
  # geom_vline(xintercept = c(quantile (rs_results$Rs_raw, c(0.025, 0.5, 0.975))),col = "red", linetype = "dotted", size = 1.5) +
  # annotate("text", x = 100, y = 0.08, label = expression("Rroot-to-Ra ratio seperated into 4 groups")
  #          , hjust = 0) +
  annotate("text", x = 100, y = 0.06, label = paste0("2.5%, 50%, 97.5% quantile (", rsg_qt[1],", ", rsg_qt[2], ", ", rsg_qt[3], ")")
           , hjust = 0, col = rgb(1,0,1)) +
  annotate("text", x = 100, y = 0.04, label = paste0("2.5%, 50%, 97.5% quantile (", rsg_qt_raw[1],", ", rsg_qt_raw[2], ", ", rsg_qt_raw[3], ")")
           , hjust = 0, col = rgb(0,1,1)) +
  ylab(expression(count)) +
  xlab(expression(Global~R[S]~"("~Pg~C~yr^{-1}~")")) -> p2

plot_grid(p1, p2, ncol=1, labels = c("(a)", "(b)")
          , hjust = -4, vjust = 4)

ggsave("outputs/Figure2.jpg", width = 6, height = 8, dpi = 300, units = "in")
```


## Other diagnose
```{r diagnose}
# check
gpp_results %>% transmute(check = (1 - Froot) / Froot - FsFr)

rs_results$Ra2 %>% min(na.rm = TRUE)
rs_results$Ra3 %>% min(na.rm = TRUE)

# check Froot (root-to-Rs ratio)
Froot %>% transmute (check = Froot + Fshoot) %>% max
Froot %>% transmute (check = Froot + Fshoot) %>% min
Froot %>% group_by(Ecosystem) %>% summarise(count = n())

Froot %>% count(IGBP2) 
Froot %>% count(Ecosystem) 
mean(Froot$Froot)
```


## Discard codes
```{r not used}
# Each row get the closese latitude and longitude values to AGB and SOC data
# AGB$Latitude[which.min(abs(61.85 - AGB$Latitude))]
# Froot %>% mutate(Latitude = round(Latitude*2, 0)/2 - 0.25, Longitude = round(Longitude*2, 0)/2-0.25) -> Froot
# there are no clear relationship between Froot and above-ground biomass or soil organic carbon (SOC)
# left_join(Froot, AGB) %>% ggplot() + aes(RASTERVALU, Froot) + geom_point()
# left_join(Froot, SOC) %>% ggplot() + aes(RASTERVALU, Froot) + geom_point()

# Each row get the closese latitude and longitude values to AGB and SOC data
# AGB$Latitude[which.min(abs(61.85 - AGB$Latitude))]
# Froot %>% mutate(Latitude = round(Latitude*2, 0)/2 - 0.25, Longitude = round(Longitude*2, 0)/2-0.25) -> Froot
# there are no clear relationship between Froot and above-ground biomass or soil organic carbon (SOC)
# left_join(Froot, AGB) %>% ggplot() + aes(RASTERVALU, Froot) + geom_point()
# left_join(Froot, SOC) %>% ggplot() + aes(RASTERVALU, Froot) + geom_point()
```