---
title: "TopdownApproach"
author: "Jinshi"
date: "May 19, 2019"
output:
  html_document: default
  word_document: default
---


## Abbreviation
* Rroot - root respiration
* Rshoot - shoot respiration
* Rs - soil respiration
* RSG - global annual soil respiration 
* Ra - autotrophic respiration (Rroot + Rshoot)
* GPP - gross primary production
* NPP - net primary production
* NEP - net ecosystem production

* Froot - Rroot / Ra (root autotrophic respiration to total autotrophic respiration fraction/ratio, calculation please see plot_froot function)
* Fshoot - Rshoot / Ra (1-Froot)
* FsFr - (1-Froot)/Froot (FsFr = Rshoot/Rroot ratio)
* RC - (root respiration / soil respiration)
* RaGPP - (Ra/GPP) ratio


## Assumptions
* __Bottom-up approach:__ 
1. We collected published global soil respiration estimates (mostly is based on bottom up field Rs measurements)
2. We then partition Rs into root respiration and belowground heterotrophic resipiration
3. We also collected published NPP estimates in the past decades from ITO et al. (2011)
4. We then collected (and calculated) Rroot/Ra ratio, according to the root respiration, GPP can be calculates as (NPP + Rroot + Rshoot)
5. We then compare this bottomup GPP with published GPP in the past decades (n=65) to see how much probability they will match

* __Top-down approach:__
1. We first collect global GPP estimates in the past decades (most is top-dow remote sensing estimates)
2. We then estimate Ra according to Ra/GPP ratio and GPP-NPP
3. Heterotrophic respiration was estimated according to Rh = NPP - HerbComsum - Fire - Csink (NEP)
4. Rs_topdown =  Rh + Rroot
5. We then compare this top-down Rs with published Rs in the past decades (n=27) to see how much probability they will match

## Questions and more analysis in the future
* How unlikely is it that we can reconcile current Rs and GPP estimates? 
E.g. what % of the distribution of each 'allows' us to match the central estimate of the other? (<30%)
* What is driving the differences between botton-up GPP and reported GPP and top-down Rs and reported Rs?
* How do we resolve this? Where is better/more data needed? 
* These questions all resolve around influence x uncertainty = importance, right? This is basically the 'Dietze graph'.
* Savanna and grassland more than 20% area, but sample numbers are too small for Frrot, RC, and Ra/Gpp ratio
* Ra < 0 how to resolve this (resolved)

## Importance/implication/benefits
* Provide another way to estimate GPP and Rs
* Provide a method to compare bottom-up GPP (start from field Rs/Froot/RC/Rroot measurements, then modelling to estimate GPP) with remote sensing or land model GPP.
* Provide a method to compare top-down Rs (Rs separated from remote sensing land model based GPP) with filed experiment based Rs (field Rs measurement, then modeling).


```{r load packages, message=FALSE, echo=FALSE}
# Constants
OUTPUT_DIR		<- "outputs/"
SEPARATOR		<- "-------------------------------------------"
DATA_DIR <- 'data'
# Create output and log folders if they do not exist
if(!file.exists(OUTPUT_DIR)) dir.create(OUTPUT_DIR)

library(kableExtra)
# Load required packages
library(data.table)
library(lubridate)
library(cowplot)
library(knitr)
library(ggpubr)
library(reshape)
library(ggplot2)
theme_set(theme_bw())
library(ggmap)
library(maps)
library(mapdata)
library(tidyr)
library(hexbin)
library(dplyr)
# Source all needed functions
source('functions.R')
```


```{r self defined functions, echo=FALSE}
# prepare Ra-GPP ratio 
prep_RaGpp <- function(RaGPP, srdb_v4) {
  # Ra/GPP
  RaGPP %>% 
    select(Ecosystem, Leaf_habit, RaGPP_ratio) %>% 
    mutate(Source = "Meta-papers") -> sdata2
  
  srdb_v4 %>% 
    # select(Study_number, Biome, Ecosystem_type, Leaf_habit, Latitude, Longitude, Ra_annual, ER, GPP, NPP) %>% 
    select(Ecosystem_type, Leaf_habit, Rs_annual, Ra_annual, Rh_annual, ER, GPP, NPP) %>% 
    filter(!is.na(GPP) | !is.na(NPP) | !is.na(ER)) %>% 
    mutate(ER = if_else(is.na(ER), GPP, ER), # assume NEP very small
           RaGPP_ratio = (ER - NPP) / GPP,
           Source = "srdb_v4") %>% 
    filter(RaGPP_ratio > 0, RaGPP_ratio < 1, !is.na(Ecosystem_type)) %>%
    select(Ecosystem_type, Leaf_habit, RaGPP_ratio, Source) %>% 
    bind_rows(sdata2)
}

# plot froot
plot_froot <- function (sdata) {
  # plot Froot (root respiration/aototrophic respiration ratio) by ecosystem type
  Fl_plot <- ggplot(sdata, aes(x = IGBP2, y=Froot)) + geom_violin() +
    geom_jitter(shape = 16, position = position_jitter(0.2), col = "gray") +
    geom_boxplot(width = 0.1) +
    ylab("Froot") +
    # stat_summary(fun.y=median, geom="point", size=2, color="red") +
    scale_x_discrete(limits = c("DF", "EF", "MF", "Other"),
                     labels = c("DF (n=13)","EF (n=91)", "MF (n=7)", "Other (n=11)") ) +
    theme(axis.title.x=element_blank()) +
    ylab("Rroot-Ra ratio")
  
  print(Fl_plot)
}
```


```{r preliminaries, echo=FALSE}
# message=FALSE, include=FALSE, echo=FALSE, cache=TRUE
# Set chunks defaults; these options will be applied to all subsequent chunks
knitr::opts_chunk$set(results = 'hide', message = TRUE, include = TRUE, 
                      echo = FALSE, warning = FALSE,
                      fig.height = 4, fig.width = 8, cache = TRUE)
```


## Methods

```{r load data}
# input data
srdb_v4 <- read.csv(file.path(DATA_DIR, 'srdbv4.csv')) # SRDB_V4 data
FlFsFr <- read.csv(file.path(DATA_DIR, 'FlFsFrSummary.csv')) # Froot (Rroot/Rs) and Fshoot (Rshoot/Rs) from 39 papers
GPP <- read.csv(file.path(DATA_DIR, 'GlobalGPP_Sum.csv')) # 65 GPP estimates from published paper
GPP %>% filter(Year > 1980) -> GPP
NPP <- read.csv(file.path(DATA_DIR, 'ITONPP.CSV')) # 251 NPP estimates from ITO (2011)
GlobalRs <- read.csv(file.path(DATA_DIR, 'GlobalRs.csv')) # 27 estimates of global Rs
GlobalRs %>% filter(Year > 1980) -> GlobalRs
# (Ra/GPP) Piao, Shilong, et al. "Forest annual carbon cost: A global‐scale analysis of autotrophic respiration." Ecology 91.3 (2010): 652-661.
RaGPP <- read.csv(file.path(DATA_DIR, 'RaGPP.csv')) 
IGBP <- read.table(file.path(DATA_DIR, 'IGBP.txt'), header = TRUE, sep = ",")
# sort(unique(IGBP$IGBP2001Pr))
# There are 16 IGBP groups
IGBP %>% 
  mutate(IGBP = case_when(IGBP2001Pr == 0 ~ "WAT",
                          IGBP2001Pr == 1 ~ "ENF",
                          IGBP2001Pr == 2 ~ "EBF",
                          IGBP2001Pr == 3 ~ "DNF",
                          IGBP2001Pr == 4 ~ "DBF",
                          IGBP2001Pr == 5 ~ "MF",
                          IGBP2001Pr == 6 ~ "CSH",
                          IGBP2001Pr == 7 ~ "OSH",
                          IGBP2001Pr == 8 ~ "WSA",
                          IGBP2001Pr == 9 ~ "SAV",
                          IGBP2001Pr == 10 ~ "GRA",
                          IGBP2001Pr == 11 ~ "WET",
                          IGBP2001Pr == 12 ~ "CRO",
                          IGBP2001Pr == 13 ~ "URB",
                          IGBP2001Pr == 14 ~ "CVM",
                          IGBP2001Pr == 15 ~ "SNO",
                          IGBP2001Pr == 16 ~ "BSV",
                          TRUE ~ "UNC" ) ) -> IGBP
# BBL: Interesting; I didn't know case_when! Aanother way to do this would be 
# IGBP_list <- c("WAT", "ENF", ...)
# IGBP$IGBP <- IGBP_list[IGBP$IGBP2001Pr]
# 16 IGBP group into 10 Ecosystems
IGBP %>% 
  mutate (Ecosystem = case_when(IGBP2001Pr == 0 ~ "Water",
                                IGBP2001Pr == 1 ~ "Forest",
                                IGBP2001Pr == 2 ~ "Forest",
                                IGBP2001Pr == 3 ~ "Forest",
                                IGBP2001Pr == 4 ~ "Forest",
                                IGBP2001Pr == 5 ~ "Forest",
                                IGBP2001Pr == 6 ~ "Shrubland",
                                IGBP2001Pr == 7 ~ "Shrubland",
                                IGBP2001Pr == 8 ~ "Savanna",
                                IGBP2001Pr == 9 ~ "Savanna",
                                IGBP2001Pr == 10 ~ "Grassland",
                                IGBP2001Pr == 11 ~ "Wetland",
                                IGBP2001Pr == 12 ~ "Agriculture",
                                IGBP2001Pr == 13 ~ "Urbland",
                                IGBP2001Pr == 14 ~ "Agriculture",
                                IGBP2001Pr == 15 ~ "Snow",
                                IGBP2001Pr == 16 ~ "Desert",
                                TRUE ~ "Unclassified")) -> IGBP

IGBP %>% 
  mutate(IGBP_group = case_when(IGBP2001Pr == 0 ~ "WAT",
                                IGBP2001Pr == 1 ~ "EF",
                                IGBP2001Pr == 2 ~ "EF",
                                IGBP2001Pr == 3 ~ "DF",
                                IGBP2001Pr == 4 ~ "DF",
                                IGBP2001Pr == 5 ~ "MF",
                                IGBP2001Pr == 6 ~ "SH",
                                IGBP2001Pr == 7 ~ "SH",
                                IGBP2001Pr == 8 ~ "GRA",
                                IGBP2001Pr == 9 ~ "GRA",
                                IGBP2001Pr == 10 ~ "GRA",
                                IGBP2001Pr == 11 ~ "WET",
                                IGBP2001Pr == 12 ~ "CRO",
                                IGBP2001Pr == 13 ~ "URB",
                                IGBP2001Pr == 14 ~ "CRO",
                                IGBP2001Pr == 15 ~ "SNO",
                                IGBP2001Pr == 16 ~ "BSV",
                                TRUE ~ "UNC" ) ) -> IGBP

# DGRsD <- read.csv(file.path(DATA_DIR, 'DGRsD.csv'), header = TRUE)
AGB <- read.table(file.path(DATA_DIR, 'GLC2000_Point.txt'), header = TRUE, sep = ",")
AGB %>% filter(RASTERVALU != -9999) -> AGB

SOC <- read.table(file.path(DATA_DIR, 'SOC_Point.txt'), header = TRUE, sep = ",")

srdb_v4 %>% 
  select(Study_number, Biome, Ecosystem_type, Leaf_habit, 
         Latitude, Longitude, RC_annual, Ra_annual, Rh_annual) %>% 
  filter(RC_annual > 0, RC_annual < 1) -> sub_srdb

# calculate Froot
Froot <- cal_Froot(FlFsFr, srdb_v4) 
# class(Froot$IGBP)
# check whether Froot + Fshoot = 1

Froot %>% mutate(IGBP2 = case_when( 
  IGBP %in% c('DBF', 'Deciduous') & Ecosystem == 'Forest' ~ "DF",
  IGBP %in% c('EBF', 'ENF', 'Evergreen') & Ecosystem == 'Forest' ~ "EF",
  IGBP %in% c('MF', 'Mixed') & Ecosystem == 'Forest' ~ "MF",
  TRUE ~ "Other"
)) -> Froot
```


## Sites spatial distribution
```{r plot sites spatial distribution, echo=FALSE, warning=FALSE}
# Figure. Sites distribution -- FlFsFr sites and RC/HC sites
plot_sites(srdb_v4, FlFsFr)
```


## Ratio by vegetation types {.tabset .tabset-fade .tabset-pills}
### Plot GPP
```{r plot GPP and RSG, fig.height = 6, fig.width = 8}
# panel (a). all 65 GPP estimates
# panel (b). GPP vs year, the red dots in panel c are calculated trend from panel b (increase rate = 0.2 if outliers removed, increase rate = 0.42 if outliers kept)
# panel (c). 16 GPP estimates reported GPP increase trend
plot_GPP(GPP, GlobalRs)
ggsave("outputs/FigureS1.jpg", width = 8, height = 6, dpi = 300, units = "in")

# A couple of simpler figures for Ben's Japan presentation
library(ggExtra)
p_gpp <- ggplot(GPP, aes(Year, GPP)) + geom_point() + 
  geom_errorbar(aes(ymin = GPP - SD, ymax = GPP + SD), na.rm = TRUE) + 
  geom_smooth(method = "lm") + theme_cowplot() +
  ylab(expression(GPP~(Pg~C~yr^-1)))
p_gpp <- ggMarginal(p_gpp, type = "densigram", margins = "y", fill = "red")
ggsave("outputs/p_gpp.png", plot = p_gpp, width = 8, height = 5)

p_rs <- ggplot(GlobalRs, aes(Year_published, Rs)) + geom_point() + 
  geom_errorbar(aes(ymin = Rs - SD, ymax = Rs + SD), na.rm = TRUE) + 
  geom_smooth(method = "lm") + theme_cowplot() +
  ylab(expression(R[S]~(Pg~C~yr^-1)))
p_rs <- ggMarginal(p_rs, type = "densigram", margins = "y", fill = "red")
ggsave("outputs/p_rs.png", plot = p_rs, width = 8, height = 5)

```


### Plot Rroot-Rs ratio

```{r plot RC, fig.height=6, fig.width=8}
# Plot global Rs, NPP, Fire-burned Carbon, and RC by ecosystem (root respiration to soil respiration ratio)
# more details please see function plot_Rroot_Rs()
# prepare data first
sub_srdb %>% 
  select(Leaf_habit) %>% 
  group_by(Leaf_habit) %>% 
  summarise(count = n())

sub_srdb %>% 
  mutate(IGBP2 = case_when( Ecosystem_type == "Agriculture" ~ "Agriculture",
                            Ecosystem_type == "Grassland" ~ "Grassland",
                            Ecosystem_type %in% c("Savanna", "Shrubland") ~ "Shrubland",
                            Ecosystem_type == "Forest" & Leaf_habit == 'Deciduous' ~ "DF",
                            Ecosystem_type == "Forest" & Leaf_habit == 'Evergreen' ~ "EF",
                            Ecosystem_type == "Forest" & Leaf_habit == 'Mixed' ~ "Mixed",
                            TRUE ~ "Other" )) -> sub_srdb
sub_srdb %>% 
  group_by(Ecosystem_type) %>% 
  summarise(count = n())

plot_Rroot_Rs (GlobalRs, sub_srdb, NPP) 
ggsave("outputs/FigureS2.jpg", width = 8, height = 6, dpi = 300, units = "in")
```

### Plot Rroot-Ra ratio
```{r plot Rroot-Ra ratio, fig.height = 4, fig.width = 6}
# potential issue: only two points from grassland
# no point from tundra 
plot_froot (Froot)
ggsave("outputs/FigureS3.jpg", width = 6, height = 4, dpi = 300, units = "in")
```


### Plot Ra to GPP ratio
```{r plot Ra-GPP ratio}
# only one point from tundra
var_RaGpp <- prep_RaGpp(RaGPP, srdb_v4)
var_RaGpp %>% count(Ecosystem)
var_RaGpp %>% count(Leaf_habit)

var_RaGpp %>% 
  mutate(IGBP2 = case_when( Ecosystem %in% c( "Crop", "Tundra", "Wetland", "Agriculture" ) ~ "Other",
                            Ecosystem %in% c("Grassland", "Savanna") ~ "Grassland",
                            Ecosystem == "Forest" & Leaf_habit == "Deciduous" ~ "Deciduous",
                            Ecosystem == "Forest" & Leaf_habit == "Evergreen" ~ "Evergreen",
                            Ecosystem == "Forest" & Leaf_habit == "Forest" ~ "Forest",
                            Ecosystem == "Forest" & Leaf_habit == "Mixed" ~ "Forest",
                            TRUE ~ "NOTA" )) -> var_RaGpp

var_RaGpp %>% count(IGBP2)
var_RaGpp %>% count(Ecosystem)

plot_RaGPP(var_RaGpp)
ggsave("outputs/FigureS4.jpg", width = 6, height = 4, dpi = 300, units = "in")

var_RaGpp %>% 
  group_by(Ecosystem) %>% 
  summarise( count = n() ) %>%
  mutate(percent = count / sum(count)) %>%
  arrange(desc(Ecosystem))
```


```{r calculate area percentage of each vegetation type}
# IGBP area by type
IGBP %>% count(IGBP_group)

IGBP %>% filter(IGBP_group != "WAT") %>% 
  group_by(IGBP_group) %>% 
  summarise(count = n()) %>% 
  mutate(percent = count / sum(count)) -> IGBP_summary

ag_area <- IGBP_summary[IGBP_summary$IGBP_group == "CRO",]$percent
df_area <- IGBP_summary[IGBP_summary$IGBP_group == "DF",]$percent
ef_area <- IGBP_summary[IGBP_summary$IGBP_group == "EF",]$percent
mf_area <- IGBP_summary[IGBP_summary$IGBP_group == "MF",]$percent
gra_area <- IGBP_summary[IGBP_summary$IGBP_group == "GRA",]$percent
shr_area <- IGBP_summary[IGBP_summary$IGBP_group == "SH",]$percent
```


```{r prepare variables for boostrapping}
# 1. prepare RC (Rroot-to-Rs ratio data) **********
# get subset data for Bottom-up approache to estimate GPP, separate into 7 groups
var_RC <- sub_srdb %>% 
  select(RC_annual, Ecosystem_type, IGBP2) %>% 
  filter(RC_annual > 0.01 & RC_annual < 0.99 & !is.na(RC_annual)) # dataset contain RC records
var_for_rc <- var_RC %>% filter (Ecosystem_type == "Forest")
# forest further separate into 3 groups
var_df_rc <- var_RC %>% filter (IGBP2 == "DF")
var_ef_rc <- var_RC %>% filter (IGBP2 == "EF")
var_mf_rc <- var_RC %>% filter (IGBP2 == "Mixed")
var_agr_rc <- var_RC %>% filter (Ecosystem_type == "Agriculture")
var_gra_rc <- var_RC %>% filter (Ecosystem_type == "Grassland")
var_shr_rc <- var_RC %>% filter (Ecosystem_type %in% c("Shrubland", "Savanna"))
# Other groups
var_other_rc <- var_RC %>% filter (Ecosystem_type %in% c("Desert", "Wetland"))
var_rest_rc <- var_RC # for those ecosystem (e.g., tundra) has no data, use all from RC dataset

# 2. prepare Froot (Rroot-to-Ra ratio data) **********
var_agr_froot <- Froot[Froot$Ecosystem == "Agriculture",]$Froot # samples of Froot for agriculture
var_for_froot <- Froot[Froot$Ecosystem == "Forest",]$Froot # samples of Froot for forest
var_gra_froot <- Froot[Froot$Ecosystem == "Grassland",]$Froot # samples of Froot for grassland
var_sav_froot <- Froot[Froot$Ecosystem == "Savanna",]$Froot # samples of Froot for savanna
var_wet_froot <- Froot[Froot$Ecosystem == "Wetland",]$Froot # samples of Froot for agriculture
var_rest_froot <- Froot$Froot # for those ecosystems has no data, use all data from Froot
# only separate into ef and the rest
var_ef_froot <- Froot[Froot$IGBP2 == "EF",]$Froot # Froot data for ef
var_df_froot <- Froot[Froot$IGBP2 == "DF",]$Froot # Froot data for df
var_mf_froot <- Froot[Froot$IGBP2 == "MF",]$Froot # Froot data for mf
var_other_froot <- Froot[Froot$IGBP2 == "Other",]$Froot # samples of Froot for non-ef
```


```{r bootstrap-creation}
# Bootstrap settings
N_SAMPLES <- 50000 # Take 50000 samples
QUANTILES <- c(0.025, 0.5, 0.975)  # Generate mean at 95% CI
set.seed(20190827)

# Use a short function to do this twice
replace_SD_and_generate <- function(m, s, n_samples = N_SAMPLES) {
  stopifnot(length(m) == length(s))
  # Alternately for first part...see #9
  # CV <- s / m
  # CV <- replace_na(CV, median(CV, na.rm = TRUE))
  # s <- if_else(is.na(s), m * CV, s)
  
  # Generate and compute mean
  empty_s <- is.na(s)
  if(sum(empty_s)) { # entries with no s.d. get a single, collective draw
    raw <- sample(m[which(empty_s)], size = n_samples, replace = TRUE)
    x <- 1
  } else {
    raw <- rep(0.0, n_samples)
    x <- 0
  }
  # Others are drawn individually
  for(i in which(!empty_s)) {
    raw <- raw + rnorm(n = n_samples, mean = m[i], sd = s[i])
    x <- x + 1
  }
  raw / x
}

bootstrap_data <- tibble(Rs_raw = replace_SD_and_generate(GlobalRs$Rs, GlobalRs$SD),
                         GPP_raw = replace_SD_and_generate(GPP$GPP, GPP$SD))
```


## Bootstrapping GPP
```{r gpp-bootstrap}
# Bottom-up approach to estimate GPP
# Rroot = Rs * RC (Rroot to Rs ratio, from srdb)
# Froot = Rroot / Ra (root autotrophic respiration to total autotrophic respiration fraction/ratio, calculation please see plot_froot function)
# Fshoot = 1-Froot
# FsFr = (1-Froot)/Froot (FsFr = Rshoot/Rroot ratio)
# Ra = Rroot + Rshoot
# GPP = NPP + Rroot + Rshoot

bootstrap_data %>% 
  mutate(
    # Step 1. Prepare NPP, boosting from ITO paper
    NPP = sample(NPP$NPP, N_SAMPLES, replace = TRUE),
    
    # Step 2. sampling Rroot-to Rs ratio (RC)
    Rc_agr = sample(var_agr_rc$RC_annual, N_SAMPLES, replace = TRUE),
    # forest further separate into 3 groups
    Rc_df = sample(var_df_rc$RC_annual, N_SAMPLES, replace = TRUE),
    Rc_ef = sample(var_ef_rc$RC_annual, N_SAMPLES, replace = TRUE),
    Rc_mf = sample(var_mf_rc$RC_annual, N_SAMPLES, replace = TRUE), 
    # other ecosystems
    Rc_gra = sample(var_gra_rc$RC_annual, N_SAMPLES, replace = TRUE),
    Rc_shr = sample(var_shr_rc$RC_annual, N_SAMPLES, replace = TRUE),
    # for the rest of ecosystems we use samples from all RC records rather that from other groups
    Rc_rest = sample(var_rest_rc$RC_annual, N_SAMPLES, replace = TRUE), 
    
    # Step 3. Sampling Rroot-to-Ra ratio (Froot)
    Froot_df = sample(var_df_froot, N_SAMPLES, replace = TRUE),
    Froot_ef = sample(var_ef_froot, N_SAMPLES, replace = TRUE),
    Froot_mf = sample(var_mf_froot, N_SAMPLES, replace = TRUE),
    # for the rest of ecosystems we used samples from all RC records rather from other groups
    Froot_rest = sample(var_rest_froot, N_SAMPLES, replace = TRUE),
    
    # Step 4. calculation ********************
    # Rc = replicate( n_rep, var_RC$RC_annual, N_SAMPLES, replace = TRUE) %>% mean() ), # scenario 0: Rc not separate into different ecosystems, not used
    Rc = sample(var_RC$RC_annual, N_SAMPLES, replace = TRUE),
    # Rc = Rc_df*df_area + Rc_ef*ef_area + Rc_mf*mf_area + Rc_rest*1-(df_area + ef_area + mf_area), 
    
    # Scenario 2: Rc separate into 4 groups: agriculture, forest, grassland, and rest (all other ecosystems with obs<40, take samples from all RC records)
    # only separate into four groups, agriculture, forest, grassland and the rest
    # 1-(0.191 + 0.214 + 0.153 + 0.268)
    Rc2 = (Rc_agr * ag_area + Rc_df * df_area + Rc_ef * ef_area + Rc_mf * mf_area + 
             Rc_gra * gra_area + Rc_shr * shr_area 
           + Rc_rest * (1 - ag_area - df_area - ef_area-mf_area - gra_area - shr_area)), 
    # mean(sample(var_for$RC_annual, N_SAMPLES, replace = TRUE))
    # sample(var_df$RC_annual, N_SAMPLES, replace = TRUE) %>% mean
    # sample(var_ef$RC_annual, N_SAMPLES, replace = TRUE) %>% mean
    # sample(var_mf$RC_annual, N_SAMPLES, replace = TRUE) %>% mean
    
    # scenario 1: Froot not separate into different ecosystems
    Froot = sample(Froot$Froot, N_SAMPLES, replace = TRUE) ,
    
    FsFr = (1 - Froot) / Froot, # calculate Froot to Fshoot ratio of scenario 1
    # Bottom up estimate of GPP, scenario 1
    Rroot = Rs_raw * Rc, # root respiration
    Rshoot = Rroot * FsFr, # shoot respiration
    
    # Scenario 2: separate into 2 ecosystems and weighted by area
    # Froot %>% count(Ecosystem)
    Froot2 = Froot_df * df_area + Froot_ef * ef_area + Froot_mf * mf_area + 
      Froot_rest * (1 - df_area - ef_area - mf_area),
    FsFr2 = (1 - Froot2) / Froot2, # Froot to Fshoot ratio of scenario 2
    # Bottom up estimate of GPP, scenario 2
    Rroot2 = Rs_raw * Rc2,
    Rshoot2 = Rroot2 * FsFr2,
    
    GPP = NPP + Rroot + Rshoot, # scenario 1: GPP
    GPP2 = NPP + Rroot2 + Rshoot2 # scenario 2: GPP2
  ) ->
  gpp_results
```


## GPP results {.tabset .tabset-fade .tabset-pills}
### Scenario 1 - ratio __not__ weighted by ecosystem
```{r plot GPP comparison}
# hist(gpp_results$GPP) + abline (v = quantile (gpp_results$GPP, QUANTILES), col = "red", lty = "dashed")
# mean(gpp_results$GPP)
# max(gpp_results$GPP)

# scenario 1
gpp_qt <- quantile(gpp_results$GPP, QUANTILES) %>% round(0)
gpp_qt_raw <- quantile(gpp_results$GPP_raw, QUANTILES) %>% round(0)

gpp_results %>%
  select(`GPP implied by Rs` = GPP, `Top-down GPP` = GPP_raw) %>% 
  gather(GPP_type) ->
  gpp_figure_data

gpp_figure_data %>% 
  ggplot(aes(value, fill = GPP_type)) +
  geom_density(stat = 'density', alpha = 0.5) +
  theme(legend.position = c(0.75, 0.75)) +
  coord_cartesian(xlim = c(25, 350)) +
  scale_fill_discrete("") + ylab("Density") +
  xlab(expression(GPP~(Pg~C~yr^{-1}))) ->
  gpp_base_figure

# We're adding same annotations to each figure so use reusable functions
make_label <- function(qt_data) {
  paste0("Mean and 95% CI: ", qt_data[2], " (", qt_data[1], ", ", qt_data[3], ")")
}
annotations <- function(figure, types, x_qt, x_qt_raw, xpos) {
  ann_dat <- tibble(Type = types,
                    GPP_type = NA, Rs_type = NA,
                    x = xpos, y = c(0.035, 0.025),
                    label = c(make_label(x_qt), make_label(x_qt_raw)))
  figure + geom_text(data = ann_dat,
                     aes(x = x, y = y, label = label, color = Type),
                     hjust = 0, show.legend = FALSE)
}

p1_gpp <- annotations(gpp_base_figure, 
                      unique(gpp_figure_data$GPP_type), gpp_qt, gpp_qt_raw, 150)
print(p1_gpp)
ggsave("outputs/Figure2_gpp.jpg", width = 8, height = 4, dpi = 300, units = "in")
```

### Scenario 2 - ratio __weighted__ by ecosystem
```{r}
# scenario 2
gpp_qt <- quantile(gpp_results$GPP2, QUANTILES) %>% round(0)
gpp_qt_raw <- quantile(gpp_results$GPP_raw, QUANTILES) %>% round(0)

gpp_results %>%
  select(`GPP implied by Rs` = GPP2, `Top-down GPP` = GPP_raw) %>% 
  gather(GPP_type) ->
  gpp_figure2_data

p1_gpp2 <- annotations(gpp_base_figure %+% gpp_figure2_data,
                       unique(gpp_figure2_data$GPP_type), gpp_qt, gpp_qt_raw, 150)
print(p1_gpp2)
```

### Scenario 3 - SRDB Rs:GPP ratios
```{r Simple SRDB Rs:GPP ratios}
var_Rs_to_GPP <- with(srdb_v4, Rs_annual / GPP) %>% na.omit

bootstrap_data %>% 
  mutate(Rs_to_GPP = sample(var_Rs_to_GPP, N_SAMPLES, replace = TRUE),
         Rs = GPP_raw * Rs_to_GPP,
         GPP = Rs_raw / Rs_to_GPP) ->
  simple_ratio_results

gpp_qt <- quantile(simple_ratio_results$GPP, QUANTILES) %>% round(0)
gpp_qt_raw <- quantile(simple_ratio_results$GPP_raw, QUANTILES) %>% round(0)

simple_ratio_results %>%
  select(`GPP implied by SRDB Rs:GPP` = GPP, `Top-down GPP` = GPP_raw) %>% 
  gather(GPP_type) ->
  gpp_simple_figure_data

gpp_simple_figure_data %>% 
  ggplot(aes(value, fill = GPP_type)) +
  geom_density(stat = 'density', alpha = 0.5) +
  theme(legend.position = c(0.75, 0.75)) +
  coord_cartesian(xlim = c(25, 350)) +
  scale_fill_discrete("") + ylab("Density") +
  xlab(expression(GPP~(Pg~C~yr^{-1}))) ->
  gpp_simple_figure

p1_gpp3 <- annotations(gpp_simple_figure,
                       unique(gpp_simple_figure_data$GPP_type), 
                       gpp_qt, gpp_qt_raw, 150)
print(p1_gpp3)
```


### GPP vs components relationships
```{r}
gpp_results %>% ggplot() + aes(Rc) + 
  # geom_histogram(col = "black", fill = "gray", bins = 30) + 
  geom_density(stat = 'density', col = "gray", fill = rgb(1, 0, 1, alpha = 0.5) )

gpp_results %>% ggplot() + aes(Rc2) + 
  # geom_histogram(col = "black", fill = "gray", bins = 30) + 
  geom_density(stat = 'density', col = "gray", fill = rgb(1, 0, 1, alpha = 0.5) )

gpp_results %>% ggplot() + aes(Rroot) + 
  # geom_histogram(col = "black", fill = "gray", bins = 30) + 
  geom_density(stat = 'density', col = "gray", fill = rgb(1, 0, 1, alpha = 0.5) )

gpp_results %>% ggplot() + aes(Rroot2) + 
  # geom_histogram(col = "black", fill = "gray", bins = 30) + 
  geom_density(stat = 'density', col = "gray", fill = rgb(1, 0, 1, alpha = 0.5) )

# GPP ~ NPP
ggplot(gpp_results) +
  aes(x = NPP, y = GPP2) +
  geom_hex() +
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") + 
  xlab(expression(NPP~"("~Pg~C~yr^{-1}~")")) +
  ylab(expression(GPP~"("~Pg~C~yr^{-1}~")"))

# GPP ~ Rs
ggplot(gpp_results) +
  aes(x = gpp_results$Rs_raw, y = GPP2) +
  geom_hex() +
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") +
  xlab(expression(R[S]~"("~Pg~C~yr^{-1}~")")) +
  ylab(expression(GPP~"("~Pg~C~yr^{-1}~")"))

# GPP ~ Rc
ggplot(gpp_results) +
  aes(x = Rc, y = GPP2) +
  geom_hex() +
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") +
  xlab(expression(RC)) +
  ylab(expression(GPP~"("~Pg~C~yr^{-1}~")"))

# GPP ~ Froot
ggplot(gpp_results) +
  aes(x = Froot, y = GPP2) +
  geom_hex() +
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") +
  xlab(expression(Froot)) +
  ylab(expression(GPP~"("~Pg~C~yr^{-1}~")"))

# GPP ~ Fshoot
ggplot(gpp_results) +
  aes(x = 1-Froot, y = GPP2) +
  geom_hex() +
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") +
  xlab(expression(Fshoot)) +
  ylab(expression(GPP~"("~Pg~C~yr^{-1}~")"))
```

### Probability of agreement
```{r compute proportion of agreement between top-down RSG and bottom-up GPP}
prob_agreement <- function(x, raw_quantiles) {
  # assuming x follows a normal distribution, we can calculate the probability
  # of it overlapping the distribution implied by the raw quantiles
  mu <- mean(x)
  s <- sd(x)
  p_high <- pnorm((raw_quantiles[3] - mu) / s)
  p_low <- pnorm((raw_quantiles[1] - mu) / s)
  (p_high - p_low) %>% round(4)
}

# Quantify the probability GPP following (107, 133)
# about 29% chance GPP will between (107-133)
p_gpp1 <- prob_agreement(gpp_results$GPP, gpp_qt_raw)
p_gpp2 <- prob_agreement(gpp_results$GPP2, gpp_qt_raw)
p_gpp3 <- prob_agreement(simple_ratio_results$GPP, gpp_qt_raw)
```

The probability of agreement in scenario 1 is ```r p_gpp1 * 100``` %. 

The probability of agreement in scenario 2 is ```r p_gpp2 * 100``` %.

The probability of agreement in scenario 3 is ```r p_gpp3 * 100``` %.


```{r prepare variables for RSG boostrapping}
# Froot used the same data as bottom-up approach
var_Fire <- c(2, 3.5, 7.3, 4, 5.1, 2.02, 2.71, 3.02, 2.08)
# prepare RaGPP ratio data
var_RaGpp %>% count(IGBP2)
var_RaGpp_df = var_RaGpp[var_RaGpp$IGBP2 == "Deciduous", ]$RaGPP_ratio
var_RaGpp_ef = var_RaGpp[var_RaGpp$IGBP2 == "Evergreen", ]$RaGPP_ratio
# mix forest are from all forest due to lack of data
var_RaGpp_mf = var_RaGpp[var_RaGpp$Ecosystem == "Forest", ]$RaGPP_ratio 
var_RaGpp_gra = var_RaGpp[var_RaGpp$Ecosystem == "Grassland", ]$RaGPP_ratio
```

## Bootstrapping RSG
```{r rs-bootstrap}
# topdowm approach to estimate Rs
# Ra = GPP * Ra_GPP_ratio (calculated based on data from a summary paper and from srdb, see plot_RaGPP function)
# Ra = GPP - NPP (two approaches yield very similar results, we thus used the average of two estimates)
bootstrap_data %>% 
  mutate(
    # Step 1. prepare other components, NPP and NPP consumed by fire etc.
    NPP = sample(NPP$NPP, N_SAMPLES, replace = TRUE),
    HerbComsum = rnorm(n = N_SAMPLES, mean = 2.2, sd = 0.2),
    Fire = sample(var_Fire, N_SAMPLES, replace = TRUE),
    sink = rnorm(n = N_SAMPLES, mean = 2.10, sd = 0.28),
    
    # Step 2. Ra-to-GPP ratio boosting
    # RaGpp = sample(var_RaGpp$RaGPP_ratio, N_SAMPLES, replace = TRUE),
    RaGpp_df = sample(var_RaGpp_df, N_SAMPLES, replace = TRUE),
    RaGpp_ef = sample(var_RaGpp_ef, N_SAMPLES, replace = TRUE),
    RaGpp_mf = sample(var_RaGpp_mf, N_SAMPLES, replace = TRUE),
    RaGpp_gra = sample(var_RaGpp_gra, N_SAMPLES, replace = TRUE),
    # for the rest of ecosystems, we use samples from all Froot rather than Other group
    RaGpp_rest = sample(var_RaGpp$RaGPP_ratio, N_SAMPLES, replace = TRUE),
    
    # Step 3. Sampling Rroot-to-Ra ratio (Froot)
    Froot_df = sample(var_df_froot, N_SAMPLES, replace = TRUE),
    Froot_ef = sample(var_ef_froot, N_SAMPLES, replace = TRUE),
    Froot_mf = sample(var_mf_froot, N_SAMPLES, replace = TRUE),
    Froot_gra = sample(var_gra_froot, N_SAMPLES, replace = TRUE),
    # for the rest of ecosystems, we use samples from all Froot rather than Other group
    Froot_rest = sample(var_rest_froot, N_SAMPLES, replace = TRUE), 
    
    # Calculation **********
    # scenario 1
    RaGpp = RaGpp_rest,
    # scenario 2
    RaGpp2 = RaGpp_df * df_area + RaGpp_ef * ef_area + RaGpp_mf * mf_area + 
      RaGpp_gra * gra_area + RaGpp_rest * (1 - df_area - ef_area - mf_area - gra_area),
    
    # get Froot (Froot2 is better, from bottom-up section, see above results)
    Froot2 = Froot_df * df_area + Froot_ef * ef_area + Froot_mf * mf_area +
      Froot_rest * (1 - df_area-ef_area-mf_area),
    
    Ra1 = GPP_raw * RaGpp, # first way to calculate Ra, Ra = GPP * RaGPP
    Ra2 = GPP_raw - NPP, # second way to calculate Ra, Ra = GPP - NPP
    Ra3 = if_else(Ra2 < 0, Ra1, Ra2),
    Ra_avg1 = (Ra1 + Ra3) / 2,
    
    Rroot = Ra_avg1 * Froot2,
    Rshoot = Ra_avg1 * (1 - Froot2),
    Rs_topdown = NPP - HerbComsum - Fire - sink + Rroot,
    
    # scenario 2
    Ra4 = GPP_raw * RaGpp2, # first way to calculate Ra, Ra = GPP * RaGPP
    Ra_avg2 = (Ra4 + Ra3) / 2,
    Rroot2 = Ra_avg2 * Froot2,
    Rshoot2 = Ra_avg2 * (1 - Froot2),
    Rs_topdown2 = NPP - HerbComsum - Fire - sink + Rroot2
  ) ->
  rs_results
```

## RSG results {.tabset .tabset-fade .tabset-pills}
### Scenario 1 - ratio __not__ weighted by ecosystem
```{r plot RSG results}

# scenario 1. RaGPP ratio not separated
rsg_qt <- quantile(rs_results$Rs_topdown, QUANTILES) %>% round(0)
rsg_qt_raw <- quantile(rs_results$Rs_raw, QUANTILES) %>% round(0)

rs_results %>%
  select(`Rs implied by GPP` = Rs_topdown, `Bottom-up Rs` = Rs_raw) %>% 
  gather(Rs_type) ->
  rs_figure_data

rs_figure_data %>% 
  ggplot(aes(value, fill = Rs_type)) +
  geom_density(stat = 'density', alpha = 0.5) +
  theme(legend.position = c(0.75, 0.75)) +
  coord_cartesian(xlim = c(25, 150)) +
  # reverse the color scale so that teal (color 1) is always observations, 
  # while pink is always implied fluxes
  scale_fill_discrete("", h.start = 180) + scale_color_discrete(h.start = 180) +
  ylab("Density") +
  xlab(expression(R[S]~(Pg~C~yr^{-1}))) ->
  rs_base_figure

p2_rs <- annotations(rs_base_figure, 
                     types = unique(rs_figure_data$Rs_type), rsg_qt, rsg_qt_raw, 100)
print(p2_rs)
ggsave("outputs/Figure2_rs.jpg", width = 8, height = 4)


# Rs raw data plot
# rsg_qt <- quantile(rs_results$Rs_raw, QUANTILES) %>% round(0)
# plot_rs_raw <- ggplot(rs_results) + aes(Rs_raw) + geom_histogram(col = "black", fill = "gray", bins = 30) + 
#   geom_vline(xintercept = c(quantile (rs_results$Rs, QUANTILES)),col = "red", linetype = "dotted", size = 1.5) +
#   annotate("text", x = quantile(rs_results$Rs, 0.98), y = 5000, label = paste0("95%CI (", rsg_qt[1],", ", rsg_qt[2], ", ", rsg_qt[3], ")"), hjust = 0)

```

### Scenario 2 - ratio __weighted__ by ecosystem
```{r}
# scenario 2. RaGPP ratio separated into 2 groups
rsg_qt <- quantile(rs_results$Rs_topdown2, QUANTILES) %>% round(0)
rsg_qt_raw <- quantile(rs_results$Rs_raw, QUANTILES) %>% round(0)

rs_results %>%
  select(`Rs implied by GPP` = Rs_topdown2, `Bottom-up Rs` = Rs_raw) %>% 
  gather(Rs_type) ->
  rs_figure_data

p2_rs2 <- annotations(rs_base_figure %+% rs_figure_data, 
                      types = unique(rs_figure_data$Rs_type), rsg_qt, rsg_qt_raw, 100)
print(p2_rs2)
```

### Scenario 3 - SRDB Rs:GPP ratios
```{r}
simple_ratio_results %>%
  select(`Rs implied by SRDB Rs:GPP` = Rs, `Top-down Rs` = Rs_raw) %>% 
  gather(Rs_type) ->
  rs_simple_figure_data

rsg_qt <- quantile(simple_ratio_results$Rs, QUANTILES) %>% round(0)
rs_qt_raw <- quantile(simple_ratio_results$Rs_raw, QUANTILES) %>% round(0)

rs_simple_figure_data %>% 
  ggplot(aes(value, fill = Rs_type)) +
  geom_density(stat = 'density', alpha = 0.5) +
  theme(legend.position = c(0.75, 0.75)) +
  coord_cartesian(xlim = c(25, 350)) +
  scale_fill_discrete("") + ylab("Density") +
  xlab(expression(R[S]~(Pg~C~yr^{-1}))) ->
  rs_simple_figure

p2_rs3 <- annotations(rs_simple_figure,
                      unique(rs_simple_figure_data$Rs_type), 
                      rsg_qt, rsg_qt_raw, 150)
print(p2_rs3)
```

### RSG and components relationship
```{r}
# Ra1 ~ Ra2
ggplot(rs_results) +
  aes(x = Ra1) + geom_histogram(col = "black", fill = "gray", bins = 30) +
  geom_histogram(aes(Ra3), col = "blue", fill = "skyblue", bins = 30, alpha = 0.5) +
  xlab(expression(Ra1~or~Ra3~"("~Pg~C~yr^{-1}~")")) +
  ylab(expression(count))

# Rs ~ GPP
ggplot(rs_results) +
  aes(x = GPP_raw, y = Rs_topdown) +
  geom_hex() +
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") +
  xlab(expression(GPP~"("~Pg~C~yr^{-1}~")")) +
  ylab(expression(R[S]~"("~Pg~C~yr^{-1}~")"))

# Rs ~ NPP
ggplot(rs_results) +
  aes(x = NPP, y = Rs_topdown) +
  geom_hex() +
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") +
  xlab(expression(NPP~"("~Pg~C~yr^{-1}~")")) +
  ylab(expression(R[S]~"("~Pg~C~yr^{-1}~")"))

# Rs ~ RaGpp
ggplot(rs_results) +
  aes(x = RaGpp, y = Rs_topdown) +
  geom_hex() +
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") +
  xlab(expression("Ra/GPP"~"("~Pg~C~yr^{-1}~")")) +
  ylab(expression(R[S]~"("~Pg~C~yr^{-1}~")"))

# Rs ~ Froot
ggplot(rs_results) +
  aes(x = Froot2, y = Rs_topdown) +
  geom_hex() +
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") +
  xlab(expression(Froot)) +
  ylab(expression(R[S]~"("~Pg~C~yr^{-1}~")"))

# Rs ~ Fire
ggplot(rs_results) +
  aes(x = Fire, y = Rs_topdown) +
  geom_hex() +
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") +
  xlab(expression(Fire~burned~"("~Pg~C~yr^{-1}~")")) +
  ylab(expression(R[S]~"("~Pg~C~yr^{-1}~")"))
```

### RSG probability of agreement

```{r compute proportion of agreement between top-down RSG and bottom-up RSG}
p_Rs1 <- prob_agreement(rs_results$Rs_topdown, rsg_qt_raw)
p_Rs2 <- prob_agreement(rs_results$Rs_topdown2, rsg_qt_raw)
p_Rs3 <- prob_agreement(simple_ratio_results$Rs, rsg_qt_raw)
```

The probability of agreement in scenario 1 is ```r p_Rs1 * 100``` %.

The probability of agreement in scenario 2 is ```r p_Rs2 * 100``` %.

The probability of agreement in scenario 3 is ```r p_Rs3 * 100``` %.


## Topdown and bottomup comparison - Figure 1
```{r combine GPP and RSG comparison into figure 2, fig.height=8, fig.width=6}
# scenario 2, RC separate into 4 groups, and Froot separate into 2 groups
plot_grid(p1_gpp2, p2_rs2, ncol=1, labels = c("(a)", "(b)"), hjust = -4, vjust = 4)
ggsave("outputs/Figure2.jpg", width = 6, height = 8, dpi = 300, units = "in")
```


## Other diagnoses
```{r diagnoses}
# check
gpp_results %>% transmute(check = (1 - Froot) / Froot - FsFr)

rs_results$Ra2 %>% min(na.rm = TRUE)
rs_results$Ra3 %>% min(na.rm = TRUE)

# check Froot (root-to-Rs ratio)
Froot %>% transmute (check = Froot + Fshoot) %>% max
Froot %>% transmute (check = Froot + Fshoot) %>% min
Froot %>% group_by(Ecosystem) %>% summarise(count = n())

Froot %>% count(IGBP2) 
Froot %>% count(Ecosystem) 
mean(Froot$Froot)
```

## Site-specific ratios of GPP to Rs

Presumably, as a matter of mathematics, the ratio of global Rs to global GPP should be the linear combination (sum) of all terrestrial points Neither SRDB nor FLUXNET is an unbiased sample of the land surface, but it may be instructive to look at the Rs:GPP for three data sources:

* Global GPP estimates, comparing them against the mean Rs estimate (~85 Pg C);
* the SRDB itself, for sites at which both GPP and Rs were measured;
* and FLUXNET2015 Tier 1, using Rs data measured very close to the towers.

```{r, site-ratios}
#topdown <- expand.grid(GPP = GPP$GPP, Rs = GlobalRs$Rs)
topdown <- data.frame(GPP = GPP$GPP, Rs = mean(GlobalRs$Rs))
topdown$Source <- paste0("Global GPP\n(N = ", nrow(topdown), ")")
topdown$Type <- "Global"

srdb_v4 %>% 
  filter(!is.na(GPP), !is.na(Rs_annual), Ecosystem_state != "Managed") %>% 
  select(Rs = Rs_annual, GPP, Biome, Ecosystem_type) %>% 
  mutate(Biome = as.character(Biome), 
         Ecosystem_type = as.character(Ecosystem_type)) ->
  srdb
srdb$Source <- paste0("SRDB\n(N = ", nrow(srdb), ")")
srdb$Type <- "Observations"

read.csv("data/site-data-temp/srdb_filtered.csv", stringsAsFactors = FALSE) %>% 
  select(Biome, Ecosystem_type, FLUXNET_SITE_ID, GPP = gpp_fluxnet, Rs = Rs_annual) %>% 
  filter(!is.na(GPP), !is.na(Rs)) ->
  fluxnet
fluxnet$Source <- paste0("FLUXNET\n(N = ", nrow(fluxnet), ")")
fluxnet$Type <- "Observations"

cmip6 <- data.frame(Source = "CMIP6\n(N = 0)", Type = "Models", stringsAsFactors = FALSE)

bind_rows(topdown, srdb, fluxnet, cmip6) %>% 
  mutate(Ratio = Rs / GPP,
         Source = factor(Source, unique(Source)[c(2,3, 1,4)])) ->
  ratios

p_ratios <- ggplot(ratios, aes(Source, Ratio, color = Type)) + 
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75), size = 0.75) +
  geom_jitter(alpha = I(0.25)) + 
  ylab("Ratio of Rs to GPP") + xlab("") +
  coord_cartesian(ylim = c(0, 2))
print(p_ratios)
```



## Discarded codes
```{r not used}
# Each row gets the closest latitude and longitude values to AGB and SOC data
# AGB$Latitude[which.min(abs(61.85 - AGB$Latitude))]
# Froot %>% mutate(Latitude = round(Latitude*2, 0)/2 - 0.25, Longitude = round(Longitude*2, 0)/2-0.25) -> Froot
# there are no clear relationship between Froot and above-ground biomass or soil organic carbon (SOC)
# left_join(Froot, AGB) %>% ggplot() + aes(RASTERVALU, Froot) + geom_point()
# left_join(Froot, SOC) %>% ggplot() + aes(RASTERVALU, Froot) + geom_point()

# Each row get the closese latitude and longitude values to AGB and SOC data
# AGB$Latitude[which.min(abs(61.85 - AGB$Latitude))]
# Froot %>% mutate(Latitude = round(Latitude*2, 0)/2 - 0.25, Longitude = round(Longitude*2, 0)/2-0.25) -> Froot
# there are no clear relationship between Froot and above-ground biomass or soil organic carbon (SOC)
# left_join(Froot, AGB) %>% ggplot() + aes(RASTERVALU, Froot) + geom_point()
# left_join(Froot, SOC) %>% ggplot() + aes(RASTERVALU, Froot) + geom_point()
```
