---
title: "1A.CMIP6_v_Obs_Comparison"
output: html_notebook
---
## Objective 

The purpose of this script is to compare the Rs to GPP ratio from the CMIP6 models and the srbd and fulxnet observations. In both cases the ratio is calcualted at the points (lat/lon) where the fluxnet and srbd observations were sampled. 



## Set Up 

Set up the environement. 
```{r, warning=FALSE, message=FALSE}
# Load the required libs. 
library(dplyr)
library(tidyr)
library(ggplot2) # For some sanity plots

# The directory of where the results will be save at. 
CMIP_DIR <- here::here('cmip')
```


Import processed CMIP data, because of the size of this data set it is not committed to the GitHub Repo, it will have to mannually be moved from pic to the project directory. 
```{r}
raw_data <- read.csv(file.path(CMIP_DIR, "annual_values_LatLon.csvNEW.csv"), stringsAsFactors = FALSE) %>%  
  filter(year %in% 2005:2014)
```



## Convert Data 

The units for the cmip output data are kg m-2 s-1 but the units for the observation data are gC m-2 yr-1 

```{r}
# Mak sure that we are only looking at the columns of relevant data. 
raw_data <- raw_data[ , names(raw_data) %in% c('value', 'units', 'variable', 'year', 'Longitude', 
                                               'Latitude', 'source', 'model','experiment', 'ensemble')]
raw_data %>%  
  filter(variable != 'areacella') %>%  
  ungroup %>% 
  # Convert from s-1 to yr -1
  mutate(value = value * 60*60*24*365, 
         units = gsub(pattern = ' ', replacement = '', x = units),
         units = gsub(pattern = 's-1', replacement = 'yr-1', x = units)) %>% 
  # Mulitply by the m2 of land 
  mutate(value = value * 1000, 
         units = gsub(pattern = 'kg', replacement = 'gC', x = units)) %>% 
  tidyr::spread(variable, value) %>% 
  na.omit ->
  wide_LatLon_fluxes
```


### Check the Quality of Data 

```{r}
summary(wide_LatLon_fluxes$gpp)
```

```{r}
summary(wide_LatLon_fluxes$raRoot)
```

```{r}
summary(wide_LatLon_fluxes$rhSoil)
```

Now calculate the period average so that it is comparable with the other observations. 
```{r}
wide_LatLon_fluxes %>% 
  group_by(units, year, Longitude, Latitude, source, model, experiment, ensemble) %>%  
  summarise(CMIP6_GPP = mean(gpp),
            CMIP6_raRoot = mean(raRoot), 
            CMIP6_rhSoil = mean(rhSoil)) %>%  
  mutate(CMIP6_Rs = CMIP6_raRoot + CMIP6_rhSoil, 
         CMIP6_Ratio = CMIP6_Rs/CMIP6_GPP) %>% 
  ungroup %>% 
  na.omit -> 
  CMIP6_LatLon_ratios
```

## Compare With Obs

```{r}
# Import the Lat and Lon information 
LatLon_Obs <- bind_rows(readRDS(file.path(CMIP_DIR, 'input', 'RsGPP_srdb_LatLon.rds')) %>% 
                           mutate(source = 'srdb'), 
                         readRDS(file.path(CMIP_DIR, 'input', 'RsGPP_fluxnet_LatLon.rds')) %>% 
                           mutate(source = 'fluxnet'))
```

```{r}
CMIP6_LatLon_ratios$Longitude <- signif(CMIP6_LatLon_ratios$Longitude, 4)
CMIP6_LatLon_ratios$Latitude  <- signif(CMIP6_LatLon_ratios$Latitude, 4)
LatLon_Obs$Longitude <- signif(LatLon_Obs$Longitude, 4)
LatLon_Obs$Latitude  <- signif(LatLon_Obs$Latitude, 4)

CMIP6_LatLon_ratios %>% 
  select(-source) %>% 
  left_join(LatLon_Obs, by = c("Longitude", "Latitude")) %>% 
  mutate(Ratio = Rs / GPP) ->
  Comparison_DF
```


```{r}
ggplot(data = Comparison_DF) + 
  geom_point(aes(Ratio, CMIP6_Ratio)) +
  labs(x = 'Observation Ratio', 
       y = 'CMIP6 Ratio', 
       title = 'A Comparison of Obs and CMIP6 Rs tp GPP Ratios by Lat/Lon') +
  geom_abline(color = 'red', slope = 1, intercept = 0)
```

Hm well if if the models were more accurate we would have expected the points to be ditributed around the 1:1 line. But the point of this is not a model vs an observation model. 

```{r}
ggplot(data = Comparison_DF) +
  geom_point(aes(Ratio, CMIP6_Ratio)) + 
  facet_wrap('source', scales = 'free') + 
    labs(x = 'Observation Ratio', 
       y = 'CMIP6 Ratio', 
       title = 'A Comparison of Obs and CMIP6 Rs tp GPP Ratios by Lat/Lon') +
  geom_abline(color = 'red', slope = 1, intercept = 0)
```


```{r}
ggplot(Comparison_DF, aes(source, CMIP6_Ratio, color = model)) + 
  geom_boxplot() +
  geom_quasirandom(alpha = I(0.25)) +
  ylab(expression(Ratio~of~R[S]~to~GPP)) + xlab("") +
  facet_wrap('ensemble', scales = 'free') + 
  theme(legend.position = 'none') + 
  coord_cartesian(ylim = c(0, 1))
```



```{r}
out_file <- file.path(CMIP_DIR, 'CMIP6_LatLon.csv')

CMIP6_LatLon_ratios %>%  
  select(year, model, experiment, ensemble, year, GPP = CMIP6_GPP, Rs = CMIP6_Rs, Longitude, Latitude) %>% 
  write.csv(file = out_file, row.names = FALSE)
cat('Saving', out_file)
```

