---
title: "1A.global_RstoGPP_processing"
output: html_notebook
---

The https://github.com/kdorheim/CMIPcarbon repo was set up to process CMIP carbon variable as part of two different carbon projects. This script has two objetives, the first being to take a look at the quality of data and to process the monthly global data to obtain annual Rs:GPP ratio for the GlobalC project. 

## Set Up 

Set up the environement. 
```{r, warning=FALSE, message=FALSE}
# Load the required libs. 
library(dplyr)
library(tidyr)
library(ggplot2) # For some sanity plots

# The directory of where the results will be save at. 
CMIP_DIR <- here::here('cmip')
```


Import processed CMIP data directly from the github repo. 
```{r}
# This data may contain results that will be used in other carbon projects. Make sure that we are only processing the variables and experiment 
# results that we are intrested in. 
raw_data <- read.csv(url('https://raw.githubusercontent.com/kdorheim/CMIPcarbon/master/pic_data/mon_global_fldmean.csv'),
         stringsAsFactors = FALSE)

```

```{r}
raw_data %>% 
  filter(variable %in% c('gpp', 'raRoot', 'rhSoil')) %>% 
  filter(experiment == 'historical') ->
  global_mean
```


## Convert from a monthly rates to annual fluxes 

Make sure there are 12 months per year. 
```{r}
global_mean %>% 
  group_by(year, ensemble, experiment, model, variable, units, area, domain, area_units) %>% 
  summarise(count = n()) %>% 
  ungroup() %>% 
  filter(count == 12) %>% 
  select(ensemble, experiment, model, variable, year) -> 
  good_months 

global_mean %>% 
  inner_join(good_months, by = c("model", "variable", "experiment", "ensemble", "year"))  -> 
  global_mean_to_process
```

Aggregate to annual rate and then calculate total flux. 
```{r}
global_mean_to_process %>%
  # Calculate the annual average weighted by the number of days in a month.
  mutate(days_per_month = lubridate::days_in_month(month)) %>% 
  group_by(units, domain, variable, model, experiment, ensemble, year, area) %>%  
  summarise(value = weighted.mean(x = value, w = days_per_month, na.rm = TRUE)) %>% 
  ungroup ->
  global_yearmonmean

global_yearmonmean %>% 
  # Convert from s-1 to yr -1
  mutate(value = value * 60*60*24*365, 
         units = gsub(pattern = ' ', replacement = '', x = units),
         units = gsub(pattern = 's-1', replacement = 'yr-1', x = units)) %>% 
  # Mulitply by the m2 of land 
  mutate(value = value * area, 
         units = gsub(pattern = 'm-2', replacement = '', x = units)) %>% 
  select(-area) %>% 
  # Convert to PgC 
  mutate(value = value * 1000/1e15, 
         units = gsub(pattern = 'kg', replacement = 'PgC', x = units)) %>%  
  # Make sure that year information is numeric not just a 
  mutate(year = as.integer(year)) %>% 
  select(units, model, experiment, ensemble, year, variable, value) %>% 
  tidyr::spread(variable, value) %>% 
  na.omit() -> 
  wide_global_fluxes
```


## Quality Control 

For quality contorl puprposes let's take a look at the GPP, raRoot, and rhSoil results seperately before calculating the ratio. 

What does gpp look like? 
```{r}
ggplot(data = wide_global_fluxes) + 
  geom_line(aes(year, gpp, color = model, group = interaction(model, ensemble)))  + 
  labs(title = 'GPP', y = 'PgCyr-1')
```

What does raRoot look like? 
```{r}
ggplot(data = wide_global_fluxes) + 
  geom_line(aes(year, raRoot, color = model, group = interaction(model, ensemble)))  + 
  labs(title = 'raRoot', y = 'PgCyr-1')
```

What does rhSoil look like? 
```{r}
ggplot(data = wide_global_fluxes) + 
  geom_line(aes(year, rhSoil, color = model, group = interaction(model, ensemble)))  + 
  labs(title = 'rhSoil', y = 'PgCyr-1')
```

If the time series look reasonable then let's continue on and calcualte the Rs:GPP ratio! 


## Calculate Ratio

```{r}
wide_global_fluxes %>% 
  mutate(Rs = raRoot + rhSoil, 
         Rs_GPP_Ratio = Rs / gpp) -> 
  global_ratio
```


Let's take a look at some more plots because sanity checks are important! 
```{r}
ggplot(data = global_ratio) + 
  geom_point(aes(Rs, gpp, color = year, shape = model, group = interaction(model, ensemble))) + 
  facet_wrap('model') + 
  labs(title = 'Rs vs. GPP', 
       y = 'GPP PgCyr-1', 
       x = 'Rs PgCyr-1')
```


```{r}
ggplot(data = global_ratio) + 
  geom_line(aes(year, Rs_GPP_Ratio, color = model, group = interaction(model, ensemble))) + 
  labs(title = 'Rs:GPP', 
       y = 'Ratio of Rs to GPP (unitless)', 
       x = 'Year')
```

## Save Data 

Well the data looks pretty good! So let's save it in the format that the rest of the other GlobalC project code expects. 





```{r}
out_file <- file.path(CMIP_DIR, 'CMIP6_annual_RStoGPP.csv')

global_ratio %>%  
  select(year, model, experiment, ensemble, year, GPP = gpp, Rs) %>% 
  write.csv(file = out_file, row.names = FALSE)
cat('Saving', out_file)
```

